---
title: "Temporal dynamics of multidimensional traits and growth"
author:
  - 杨菁
date: '`r format(Sys.Date())`'
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    number_sections: false
    toc_float:
      collapsed: no
      smooth_scroll: no
editor_options: 
  chunk_output_type: console
---

<style type='text/css'>

body{ /* Normal  */
      font-size: 16px;
      font-family: 'Times New Roman'
  }
td {  /* Table  */
  font-size: 14px;
}
h1.title {
  font-size: 34px;
  color: Black;
  font-family: 'Arial';
}
h1 { /* Header 1 */
  font-size: 26px;
  color: DarkBlue;
  font-family: 'Arial';
}
h2 { /* Header 2 */
    font-size: 22px;
    font-family: 'Arial';
    color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: 'Arial';
  color: DarkBlue;
}

h4 { /* Header 4 */
  font-size: 16px;
  font-family: 'Arial';
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 12px;
}
</style>


```{r Chunk 1, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  tidy = TRUE)

# Load required libraries
required_libraries <- c(
  "ggplot2", "patchwork", "knitr", "kableExtra", "hypervolume", "snow",
  "parallel", "ggsignif", "alphahull", "RColorBrewer", "formatR", "gridExtra",
  "ggpubr", "fmsb", "tidyr", "ggsci", "TPD", "ggridges", "reshape2", "glmm.hp",
  "lmerTest", "ape", "nlme", "MuMIn", "performance", "sjPlot", "tibble","ggthemes",
  "effectsize", "cowplot", "ggcorrplot", "dplyr", "funspace", "purrr", "ggpmisc",
  "ggExtra", "officer", "eoffice", "broom.mixed", "lme4", "vegan", "ade4", "GGally")
invisible(suppressPackageStartupMessages(lapply(required_libraries, library, character.only = TRUE)))
source("visualization_corefun.R")
# Source custom functions and set font family
my_family <- "sans"

# Create output directories
output_dirs <- c("02_output_results_new", "02_output_results_new/analysis_results", "02_output_results_new/figs", "02_output_results_new/tables")
invisible(sapply(output_dirs, function(dir) dir.create(dir, showWarnings = FALSE)))

# # Set output directory names
dir_name <- "02_output_results_new/analysis_results"
f_dir_name <- "02_output_results_new/figs"
t_dir_name <- "02_output_results_new/tables"

# Traits
above_traits <- c("LA", "SLA", "LDMC", "LTO", "SSD", "Height")
fineroot_traits <- c("SRL", "RD", "RTD","SRA")
trait.names <- c(above_traits, fineroot_traits)
pca_traits <- c("Above_PC1", "Above_PC2","Above_PC3", "Below_PC1", "Below_PC2")  # PCA components

# Write selected traits to a text file
writeLines(trait.names, file.path(dir_name, "trait_name.txt"))

species <- c("D. oldhami", "L. glaber", "C. sclerophylla", "C. glauca", "S. superba", "H. acerba", "Q. chenii")
sp_abbr <- c("DO", "LG", "CS", "CG", "SS", "HA", "QC")
names(sp_abbr) <- species
growth_cols <- c(
  "RGR_H_1yr", "RGR_H_mean", "AGR_H_1yr", "AGR_H_mean",
  "RGR_BM_mean", "RGR_ABM_mean", "RGR_UBM_mean",
  "AGR_BM_mean", "AGR_ABM_mean", "AGR_UBM_mean",
  "RGR_BM_1yr", "RGR_ABM_1yr", "RGR_UBM_1yr",
  "AGR_BM_1yr", "AGR_ABM_1yr", "AGR_UBM_1yr")

my_colors <- c(
  "Environment"  = "#7C93C3",
  "Above-ground" = "#A5B68D",
  "Below-ground" = "#AB886D",
  "Trait*Trait"  = "#FFE1AF",
  "Trait*Env"    = "#C2D9FF")
```

# data preprocessing 
```{r Chunk 2,fig.width=8,fig.height=4.5}
load("01_data/processed_data/trait_growth_data.Rdata")  # object: tda (5818 seedlings)
table(tda$comp_year)
tda <- tda %>%
  mutate(
    RGR = RGR_H_1yr,
    RSR = UBM / ABM,
    # Unit conversion for LTO
    LTO = LTO * 0.0098 / (pi * 0.001)
  ) %>%
  select(
    sp, sp_abbr, Nenv, Light, Moisture, AP, Compete, growth_time, comp_year, ring.num, memo,
    RII, final_height, 
    RGR, RSR, UBM, ABM, BM,
    all_of(trait.names)
  )
table(tda$memo)

library(dplyr)

neg_growth_summary <- tda %>%
  filter(memo == "alive", BM > 0) %>%        # 只保留存活且有生物量的个体
  mutate(neg_RGR = RGR < 0) %>% 
  group_by(comp_year) %>%
  summarise(
    n_total = n(),
    n_RGR_le_0 = sum(neg_RGR, na.rm = TRUE),
    prop_RGR_le_0 = n_RGR_le_0 / n_total
  )

neg_growth_summary

```

## clean RGR
```{r}
# -----------------------------
# Diagnostics before cleaning (optional but useful)
# -----------------------------
counts_per_year <- tda %>%
  filter(memo != "S", !is.na(RGR), RGR <= 0) %>%
  group_by(comp_year) %>%
  summarise(n_ind = n(), .groups = "drop")

counts_per_year_complete_traits <- tda %>%
  filter(memo != "S", !is.na(RGR), RGR <= 0) %>%
  filter(if_all(all_of(trait.names), ~ !is.na(.))) %>%
  group_by(comp_year) %>%
  summarise(n_ind = n(), .groups = "drop")

# -----------------------------
# Clean RGR: set non-positive and outliers to NA
# -----------------------------
n_rgr_nonpos <- sum(tda$RGR <= 0, na.rm = TRUE)
tda <- tda %>%
  mutate(
    RGR = if_else(!is.na(RGR) & RGR <= 0, NA_real_, as.numeric(RGR))
  )
mean_rgr <- mean(tda$RGR, na.rm = TRUE)
sd_rgr   <- sd(tda$RGR, na.rm = TRUE)
upper <- mean_rgr + 4 * sd_rgr
lower <- mean_rgr - 4 * sd_rgr
n_rgr_outlier <- sum(tda$RGR > upper | tda$RGR < lower, na.rm = TRUE)
tda <- tda %>%
  mutate(
    RGR = if_else(!is.na(RGR) & (RGR > upper | RGR < lower), NA_real_, RGR)
  )
n_dead <- sum(tda$memo == "S", na.rm = TRUE)
n_rgr_na_not_dead <- sum(is.na(tda$RGR) & tda$memo != "S", na.rm = TRUE)

cat("RGR <= 0 set to NA:", n_rgr_nonpos, "\n")
cat("RGR outliers (±4 SD) set to NA:", n_rgr_outlier, "\n")
cat("Dead individuals (memo == 'S'):", n_dead, "\n")
cat("RGR is NA but memo != 'S':", n_rgr_na_not_dead, "\n")

# Keep only living individuals with valid RGR
tda <- tda %>%
  filter(memo != "S", !is.na(RGR), RGR > 0) ## 4171 inds
```

## prepare trait data
```{r}
ordered_cols <- c(
  "Nenv", "Light", "Moisture", "AP", "Compete", "sp", "sp_abbr", "growth_time", "comp_year", "ring.num",
  "final_height", "BM", "ABM", "RSR", "UBM", "RGR",trait.names)
indtraits_tda <- tda %>%
  select(all_of(ordered_cols)
  ) %>%
  mutate(across(all_of(trait.names), as.numeric))

# Traits to log-transform then scale (keep your original exceptions)
traits_log_scale <- setdiff(trait.names, c("LTO", "LDMC", "RD"))
traits_scale_only <- c("LTO", "LDMC", "RD")

indtraits_tda <- indtraits_tda %>%
  mutate(
    across(all_of(traits_log_scale), ~ as.numeric(scale(log10(.)))),
    across(all_of(traits_scale_only), ~ as.numeric(scale(.))))

# Standardize growth_time (optional: keep if used later)
indtraits_tda <- indtraits_tda %>%
  mutate(growth_time = as.numeric(scale(as.numeric(growth_time))))

# Extract main trait-environment data (first 25 columns)
pcatraits_tda <- na.omit(indtraits_tda) ## 3670
pcatraits_tda$sp <- factor(pcatraits_tda$sp, levels = species)
# Arrange columns in the desired order for analysis and output
pcatraits_tda <- pcatraits_tda %>% dplyr::select(all_of(ordered_cols))
table(pcatraits_tda$comp_year)

# Check functional space dimensions for aboveground and fineroot traits
funspaceDim(na.omit(pcatraits_tda[, above_traits]))
funspaceDim(na.omit(pcatraits_tda[, fineroot_traits]))
```

# Results
## Figure 2 Principal Component Analysis (PCA)
```{r Chunk 4,fig.height=8,fig.width=8}
source("visualization_corefun_new.R")
############################### Principal Component Analysis (PCA) 
# Perform PCA for aboveground traits
above_pr <- summary(princomp(pcatraits_tda[, above_traits], cor = TRUE))
pcatraits_tda[paste0("Above_PC", 1:length(above_traits))] <- above_pr$scores
pcatraits_tda$Above_PC1 <- 1 * pcatraits_tda$Above_PC1
pcatraits_tda$Above_PC2 <- 1 * pcatraits_tda$Above_PC2
# Perform PCA for fineroot traits
#froot_pr <- summary(princomp(total_scaled_tda[, fineroot_traits], cor = TRUE))
library(psych)
froot_pr <- principal(
  pcatraits_tda[, fineroot_traits],
  nfactors = length(fineroot_traits),
  rotate = "varimax",
  scores = TRUE,
  covar = TRUE)

pcatraits_tda[paste0("Below_PC", 1:length(fineroot_traits))] <- froot_pr$scores
pcatraits_tda$Below_PC1 <- -1 * pcatraits_tda$Below_PC1 
pcatraits_tda$Below_PC2 <- 1 * pcatraits_tda$Below_PC2 

above_loadings_matrix <- as.matrix(above_pr$loadings)
above_loadings_df <- data.frame(Variable = rownames(above_loadings_matrix), above_loadings_matrix)
above_loadings_df <- above_loadings_df[1:6, ]
above_loadings_df[, -1] <- round(above_loadings_df[, -1], 2)
rownames(above_loadings_df) <- paste0("PC", 1:6)
write.csv(above_loadings_df, file = "02_output_results_new/tables/above_loadings_export.csv", row.names = FALSE)

froot_loadings_matrix <- as.matrix(froot_pr$loadings)
froot_loadings_df <- data.frame(Variable = rownames(froot_loadings_matrix), froot_loadings_matrix)
froot_loadings_df <- froot_loadings_df[1:4, ]
froot_loadings_df[, -1] <- round(froot_loadings_df[, -1], 2)
rownames(froot_loadings_df) <- paste0("PC", 1:4)
write.csv(froot_loadings_df, file = "02_output_results_new/tables/froot_loadings_export.csv", row.names = FALSE)

cp_shape <- c(21,21,21)

above_pca <- Above_Comp_PCA(pcatraits_tda, above_pr, above_traits, 
                            "(a) Aboveground", c(-5.2, 5), c(-5,4.6), cp_shape) +
  theme(legend.position = "none")

above_pca2 <- Above_Comp_PCA2(pcatraits_tda, above_pr, above_traits, 
                              "(b) Aboveground", c(-5.2, 5), c(-4, 4.6), cp_shape) + 
  theme(legend.position = "none")

froot_pca <- Below_Comp_PCA(pcatraits_tda, froot_pr, fineroot_traits, 
                            "(c) Fine roots", c(-4,4.3), c(-5,5), cp_shape) + 
  theme(legend.position = "none")

above_pca_plot  <- ggMarginal(above_pca,  type = "boxplot",
                              groupColour = TRUE, groupFill = FALSE,
                              fill = "white")

above_pca_plot2 <- ggMarginal(above_pca2, type = "boxplot",
                              groupColour = TRUE, groupFill = FALSE,
                              fill = "white")

froot_pca_plot  <- ggMarginal(froot_pca,  type = "boxplot",
                              groupColour = TRUE, groupFill = FALSE,
                              fill = "white")

fig3_combined_pca_plot2 <- plot_grid(above_pca_plot, above_pca_plot2, froot_pca_plot,
                                     labels = NULL, ncol = 2)
fig3_combined_pca_plot2

ggsave(fig3_combined_pca_plot2,file="02_output_results_new/figs/fig2_pca_year.pdf",height = 7.5,width = 7.5)
```

# Trait-Growth model R2 changes over years
```{r}
pcatraits_tda1 <- pcatraits_tda
pcatraits_tda1$Above_PC1 <- scale(pcatraits_tda1$Above_PC1)
pcatraits_tda1$Above_PC2 <- scale(pcatraits_tda1$Above_PC2)
pcatraits_tda1$Above_PC3 <- scale(pcatraits_tda1$Above_PC3)
pcatraits_tda1$Below_PC1 <- scale(pcatraits_tda1$Below_PC1)
pcatraits_tda1$Below_PC2 <- scale(pcatraits_tda1$Below_PC2)
pcatraits_tda1$Light <- scale(pcatraits_tda1$Light)
pcatraits_tda1$Moisture <- scale(pcatraits_tda1$Moisture)
pcatraits_tda1$AP <- scale(pcatraits_tda1$AP)
pcatraits_tda1$growth_time <- scale(as.numeric(pcatraits_tda1$growth_time))
pcatraits_tda1$RGR_scale <- scale(pcatraits_tda1$RGR)

tda_1st <- pcatraits_tda1[pcatraits_tda1$comp_year == "1st", ]
tda_2nd <- pcatraits_tda1[pcatraits_tda1$comp_year == "2nd", ]
tda_3rd <- pcatraits_tda1[pcatraits_tda1$comp_year == "3rd", ]
```

## Part models
### Part models----Community-level
```{r}
library(lmerTest)
library(MuMIn)
library(dplyr)
library(broom.mixed) 
library(ggplot2)
library(ggpmisc)

extract_model_stats <- function(dat, t_a, t_b, e_v) {
  
  # Define full model formula (M5) including all interaction terms
  f5 <- paste("RGR ~", e_v, "+", t_a, "+", t_b, "+", 
              t_a,":",t_b, "+", 
              t_a,":",e_v, "+", 
              t_b,":",e_v, "+ (1|sp) + (1|Nenv)")
  
  # Fit the model (Removed tryCatch: Errors will stop execution for debugging)
  mod <- lmer(as.formula(f5), data = dat, REML = FALSE, 
              control = lmerControl(optimizer = "bobyqa"))
  
  # --- A. Extract Marginal R2 ---
  r2_out <- MuMIn::r.squaredGLMM(mod)
  
  if (is.matrix(r2_out)) {
    r2_val <- r2_out[1, "R2m"] 
  } else {
    r2_val <- r2_out["R2m"]
  }
  
  if (is.na(r2_val) || is.nan(r2_val)) r2_val <- NA_real_
  
  # --- B. Calculate RGR statistics for the data used in model ---
  model_frame <- model.frame(mod)
  mean_rgr <- mean(model_frame[, "RGR"], na.rm = TRUE)
  var_rgr  <- var(model_frame[, "RGR"], na.rm = TRUE)
  
  # --- C. Extract fixed effects coefficients ---
  coefs <- broom.mixed::tidy(mod, effects = "fixed") %>%
    select(term, estimate, std.error, statistic, p.value)
  
  # Map terms to logical groups
  coefs <- coefs %>%
    mutate(
      Effect_Type = case_when(
        term == "(Intercept)" ~ "Intercept",
        term == e_v ~ "Env",
        term == t_a ~ "Above",
        term == t_b ~ "Below",
        grepl(":", term) & grepl(t_a, term) & grepl(t_b, term) ~ "TT",
        grepl(":", term) & grepl(e_v, term) ~ "TE", 
        TRUE ~ "Other"
      )
    )
  
  # Append model-level statistics
  coefs$Marginal_R2 <- as.numeric(r2_val)
  coefs$Mean_RGR <- mean_rgr
  coefs$Var_RGR <- var_rgr
  
  return(coefs)
}

above_traits <- c("Above_PC1", "Above_PC2", "Above_PC3")
below_traits <- c("Below_PC1", "Below_PC2")
env_vars     <- c("Light", "Moisture", "AP", "Compete")
combinations <- expand.grid(Above = above_traits, Below = below_traits, Env = env_vars, stringsAsFactors = FALSE)
year_datasets <- list("1st" = tda_1st, "2nd" = tda_2nd, "3rd" = tda_3rd)
model_coefficients_results <- data.frame()

for (year_name in names(year_datasets)) {
  current_dat <- year_datasets[[year_name]]
  
  for (i in 1:nrow(combinations)) {
    
    curr_above <- combinations$Above[i]
    curr_below <- combinations$Below[i]
    curr_env   <- combinations$Env[i]
    
    res_df <- extract_model_stats(current_dat, curr_above, curr_below, curr_env)
    
    if (is.null(res_df)) next
    
    # Add metadata
    res_df$Year <- year_name
    res_df$Combination_ID <- i
    res_df$Trait_Above_Name <- curr_above
    res_df$Trait_Below_Name <- curr_below
    res_df$Env_Var_Name <- curr_env
    
    model_coefficients_results <- rbind(model_coefficients_results, res_df)
  }
}


```

### Part models----Species-level
```{r}
library(lmerTest)
library(MuMIn)
library(dplyr)
library(broom.mixed) 

# ==============================================================================
# 1. Function to extract stats (Species-Specific Model)
# ==============================================================================
extract_species_stats <- function(dat, t_a, t_b, e_v) {
  
  # --- Sample Size Check ---
  # Remove NAs first to check effective sample size for modeling
  # (Note: lmer will drop NAs automatically, but we check beforehand)
  valid_rows <- sum(!is.na(dat$RGR) & !is.na(dat[[t_a]]) & 
                    !is.na(dat[[t_b]]) & !is.na(dat[[e_v]]))
  
  if (valid_rows < 10) return(NULL) # Return NULL to signal skipping
  
  # Define model formula (Removed (1|sp) as we are within a single species)
  f5 <- paste("RGR ~", e_v, "+", t_a, "+", t_b, "+", 
              t_a,":",t_b, "+", 
              t_a,":",e_v, "+", 
              t_b,":",e_v, "+ (1|Nenv)")
  
  # Fit the model
  mod <- tryCatch({
    lmer(as.formula(f5), data = dat, REML = FALSE, 
         control = lmerControl(optimizer = "bobyqa"))
  }, error = function(e) NULL)
  
  if (is.null(mod)) return(NULL)
  
  # --- A. Extract Marginal R2 ---
  r2_out <- tryCatch(MuMIn::r.squaredGLMM(mod), error = function(e) NA)
  
  if (is.matrix(r2_out)) {
    r2_val <- r2_out[1, "R2m"] 
  } else if (!is.na(r2_out)) {
    r2_val <- r2_out["R2m"]
  } else {
    r2_val <- NA_real_
  }
  
  if (is.na(r2_val) || is.nan(r2_val)) r2_val <- NA_real_
  
  # --- B. Calculate RGR statistics ---
  model_frame <- model.frame(mod)
  mean_rgr <- mean(model_frame[, "RGR"], na.rm = TRUE)
  var_rgr  <- var(model_frame[, "RGR"], na.rm = TRUE)
  
  # --- C. Extract fixed effects coefficients ---
  coefs <- broom.mixed::tidy(mod, effects = "fixed") %>%
    select(term, estimate, std.error, statistic, p.value)
  
  # Map terms
  coefs <- coefs %>%
    mutate(
      Effect_Type = case_when(
        term == "(Intercept)" ~ "Intercept",
        term == e_v ~ "Env",
        term == t_a ~ "Above",
        term == t_b ~ "Below",
        grepl(":", term) & grepl(t_a, term) & grepl(t_b, term) ~ "TT",
        grepl(":", term) & grepl(e_v, term) ~ "TE", 
        TRUE ~ "Other"
      )
    )
  
  # Append stats
  coefs$Marginal_R2 <- as.numeric(r2_val)
  coefs$Mean_RGR <- mean_rgr
  coefs$Var_RGR <- var_rgr
  
  return(coefs)
}

# ==============================================================================
# 2. Prepare Loop Variables
# ==============================================================================
above_traits <- c("Above_PC1", "Above_PC2", "Above_PC3")
below_traits <- c("Below_PC1", "Below_PC2")
env_vars     <- c("Light", "Moisture", "AP", "Compete")
combinations <- expand.grid(Above = above_traits, Below = below_traits, Env = env_vars, stringsAsFactors = FALSE)

year_datasets <- list("1st" = tda_1st, "2nd" = tda_2nd, "3rd" = tda_3rd)
species_list  <- unique(tda_1st$sp) # Ensure 'sp' column exists

species_results <- data.frame()

# ==============================================================================
# 3. Main Nested Loop: Year -> Species -> Combination
# ==============================================================================
for (year_name in names(year_datasets)) {
  current_dat <- year_datasets[[year_name]]
  
  for (sp_name in species_list) {
    
    # Subset data for current species
    sp_dat <- current_dat[current_dat$sp == sp_name, ]
    
    # Check total rows before entering combination loop (Optimization)
    if (nrow(sp_dat) < 10) next 
    
    for (i in 1:nrow(combinations)) {
      
      curr_above <- combinations$Above[i]
      curr_below <- combinations$Below[i]
      curr_env   <- combinations$Env[i]
      
      # Extract stats
      res_df <- extract_species_stats(sp_dat, curr_above, curr_below, curr_env)
      
      # Handling insufficient data or model failure
      if (is.null(res_df)) {
        # Optional: create a row with NAs if you want to track failures explicitly
        # For now, we just skip
        next 
      }
      
      # Add metadata
      res_df$Year <- year_name
      res_df$Species <- sp_name  # <--- New Metadata
      res_df$Combination_ID <- i
      res_df$Trait_Above_Name <- curr_above
      res_df$Trait_Below_Name <- curr_below
      res_df$Env_Var_Name <- curr_env
      
      species_results <- rbind(species_results, res_df)
    }
  }
}

# ==============================================================================
# 4. Post-processing
# ==============================================================================
head(species_results)

final_species_coefs <- species_results
```

## Figure 3
```{r}
library(dplyr)
library(ggplot2)
library(ggpmisc)
library(patchwork)

df_community_raw <- model_coefficients_results %>%
  select(Year, Combination_ID, Marginal_R2, Mean_RGR) %>%
  distinct()

library(dplyr)

df_community_raw1 <- df_community_raw %>%
  mutate(
    Year_num = case_when(
      Year == "1st" ~ 1,
      Year == "2nd" ~ 2,
      Year == "3rd" ~ 3
    )
  )

library(lme4)
library(lmerTest)

m_year <- lmer(
  Marginal_R2 ~ Year_num + (1 | Combination_ID),
  data = df_community_raw1
)

summary(m_year)
tab_model(m_year)

df_community_summary <- df_community_raw %>%
  group_by(Year) %>%
  summarise(
    R2  = mean(Marginal_R2, na.rm = TRUE),
    SE  = sd(Marginal_R2, na.rm = TRUE) / sqrt(n()),
    RGR = mean(Mean_RGR, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Year = factor(Year, levels = c("1st", "2nd", "3rd")),
    Level = "Community"
  )

df_species_raw <- final_species_coefs %>%
  select(Year, Species, Combination_ID, Marginal_R2, Mean_RGR) %>%
  distinct() 

library(dplyr)

df_species_raw1 <- df_species_raw %>%
  mutate(
    Year_num = case_when(
      Year == "1st" ~ 1,
      Year == "2nd" ~ 2,
      Year == "3rd" ~ 3,
      TRUE ~ NA_real_
    ),
    Species = factor(Species),
    Combination_ID = factor(Combination_ID)
  )
library(lme4)
library(lmerTest)

m_sp2 <- lmer(
  Marginal_R2 ~ Year_num + (1 | Species) + (1 | Combination_ID),
  data = df_species_raw1,
  REML = TRUE
)

summary(m_sp2)


df_species_summary <- df_species_raw %>%
  group_by(Year, Species) %>%
  summarise(
    R2  = mean(Marginal_R2, na.rm = TRUE),
    # 修正关键点：这里明确计算 24 个组合的 SE
    # SE = SD(24个R2) / sqrt(24)
    SE  = sd(Marginal_R2, na.rm = TRUE) / sqrt(n_distinct(Combination_ID)), 
    RGR = mean(Mean_RGR, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Year = factor(Year, levels = c("1st", "2nd", "3rd")),
    Level = "Species" 
  )

pcatraits_tda$RGR

df_combined <- bind_rows(
  df_community_summary %>% select(Year, R2, SE, RGR, Level),
  df_species_summary %>% select(Year, R2, SE, RGR, Level)
) %>%
  mutate(Level = factor(Level, levels = c("Community", "Species")))

p_a <- ggplot(df_community_summary, aes(x = Year, y = R2)) +
  geom_errorbar(aes(ymin = R2 - SE, ymax = R2 + SE, color = Year),
                width = 0.15, size = 1) +
  geom_point(aes(color = Year), size = 3) +
  geom_text(aes(label = paste0("Mean RGR: ", sprintf("%.2f", RGR))),
            color = "black",
            vjust = -2.0,
            hjust = 0.5,
            size = 3.5, fontface = "plain") +
  scale_color_manual(values = c("1st"="#FFE7CC", "2nd"="#F8CBA6", "3rd"="#E08E6D")) +
  labs(
    title = "(a) Community-level predictive power",
    y = expression("Marginal " * italic(R)^2 * "of growth models"),
    x = "Harvest year"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.2))) +
  theme_Publication(base_size = 11) +
  theme(legend.position = "none")

p_b <- ggplot(df_species_summary, aes(x = Year, y = R2)) +
  geom_violin(aes(color = Year), fill = NA, trim = FALSE, size = 0.8) +
  geom_jitter(aes(color = Year), shape = 17, size = 2.5, alpha = 0.8, width = 0.1) +
  scale_color_manual(values = c("1st"="#FFE7CC", "2nd"="#F8CBA6", "3rd"="#E08E6D")) +
  labs(
    title = "(b) Species-level predictive power",
    y = NULL,
    x = "Harvest year"
  ) +
  theme_Publication(base_size = 11) +
  theme(legend.position = "none")

p_c <- ggplot(df_combined, aes(x = RGR, y = R2)) +
  geom_errorbar(aes(ymin = R2 - SE, ymax = R2 + SE, color = Year),
                width = 0, size = 0.8, alpha = 1) +
  geom_point(aes(color = Year, shape = Level, size = Level), alpha = 95) +
  geom_smooth(method = "lm", color = "black", se = TRUE, size = 0.8, fill = "grey80",alpha=0.6) +
  stat_poly_eq(
    aes(label = paste(after_stat(rr.label), after_stat(p.value.label), sep = "*\", \"*")),
    formula = y ~ x, parse = TRUE,
    label.x = "right", label.y = "bottom", size = 4
  ) +
  scale_color_manual(values = c("1st"="#FFE7CC", "2nd"="#F8CBA6", "3rd"="#E08E6D")) +
  scale_shape_manual(values = c("Community" = 19, "Species" = 17)) +
  scale_size_manual(values = c("Community" = 3.5, "Species" = 3)) +
  scale_x_log10(
    breaks = c(0.1, 0.15, 0.2, 0.3, 0.4, 0.5),
    labels = scales::label_number(drop0trailing = TRUE)
  ) +
  labs(
    title = "(c) Effect of growth rate (Combined levels)",
    y = expression("Marginal " * italic(R)^2 %+-% "SE"),
    x = "Mean relative growth rate (log scale)"
  ) +
  theme_Publication(base_size = 11)

final_plot <- (p_a + p_b) / p_c + plot_layout(heights = c(1, 0.9))

print(final_plot)

ggsave(
  final_plot <- (p_a + p_b) / p_c + plot_layout(heights = c(1, 0.9)),
  file = "02_output_results_new/figs/fig3_part_growth_r2.pdf",
  height = 7.5, width = 7
)
library(eoffice)
topptx(final_plot,filename = "02_output_results_new/figs/fig3_part_growth_r2.pptx",
  height = 6.8, width = 6.5
)


```

## Full model---community-level
```{r}
library(lme4)
library(lmerTest)
library(dplyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(MuMIn) # 必须加载 MuMIn 包来计算 R2

# ==============================================================================
# 1. 运行模型 (保持不变)
# ==============================================================================
formula_full <- RGR_scale ~ Light + Moisture + AP + Compete + growth_time +
  Above_PC1 + Above_PC2 + Above_PC3 + Below_PC1 + Below_PC2 +
  Above_PC1:Below_PC1 + Above_PC2:Below_PC1 + Above_PC3:Below_PC1 +
  Above_PC1:Below_PC2 + Above_PC2:Below_PC2 + Above_PC3:Below_PC2 +
  Below_PC1:Below_PC2 + Above_PC1:Above_PC2 + Above_PC1:Above_PC3 +
  Above_PC2:Above_PC3 + Above_PC1:Light + Above_PC1:Moisture + Above_PC1:AP + Above_PC1:Compete +
  Above_PC2:Light + Above_PC2:Moisture + Above_PC2:AP + Above_PC2:Compete +
  Above_PC3:Light + Above_PC3:Moisture + Above_PC3:AP + Above_PC3:Compete +
  Below_PC1:Light + Below_PC1:Moisture + Below_PC1:AP + Below_PC1:Compete +
  Below_PC2:Light + Below_PC2:Moisture + Below_PC2:AP + Below_PC2:Compete +
  (1 | sp) + (1 | Nenv)

# 确保数据已加载
lm1 <- lmer(formula_full, data = tda_1st, REML = FALSE)
lm2 <- lmer(formula_full, data = tda_2nd, REML = FALSE)
lm3 <- lmer(formula_full, data = tda_3rd, REML = FALSE)

# ==============================================================================
# 2. 目标变量顺序 (Target Order - 包含 Compete)
# ==============================================================================
target_order <- c(
  # --- Single Predictors ---
  "Above_PC1", "Above_PC2", "Above_PC3",
  "Below_PC1", "Below_PC2",
  "Light", "Moisture", "AP", "Compete", 
  
  # --- Trait x Trait ---
  "Above_PC1:Below_PC1", "Above_PC1:Below_PC2",
  "Above_PC2:Below_PC1", "Above_PC2:Below_PC2",
  "Above_PC3:Below_PC1", "Above_PC3:Below_PC2",
  "Below_PC1:Below_PC2",
  "Above_PC1:Above_PC2", "Above_PC1:Above_PC3", "Above_PC2:Above_PC3",
  
  # --- Trait x Env ---
  "Above_PC1:Light", "Above_PC1:Moisture", "Above_PC1:AP", "Above_PC1:Compete",
  "Above_PC2:Light", "Above_PC2:Moisture", "Above_PC2:AP", "Above_PC2:Compete",
  "Above_PC3:Light", "Above_PC3:Moisture", "Above_PC3:AP", "Above_PC3:Compete",
  "Below_PC1:Light", "Below_PC1:Moisture", "Below_PC1:AP", "Below_PC1:Compete",
  "Below_PC2:Light", "Below_PC2:Moisture", "Below_PC2:AP", "Below_PC2:Compete"
)

# ==============================================================================
# 3. 智能匹配函数 (get_model_data_v3 - 保持不变)
# ==============================================================================
get_model_data <- function(model, target_vars) {
  summ <- summary(model)
  coefs <- as.data.frame(summ$coefficients)
  coefs$RawName <- rownames(coefs)
  matched_data <- list()
  
  for (target in target_vars) {
    found <- FALSE
    # 策略 1: 精确匹配
    if (target %in% coefs$RawName) {
      row <- coefs[coefs$RawName == target, ]
      row$DisplayName <- target
      matched_data[[target]] <- row
      found <- TRUE
    } 
    # 策略 2: 反转匹配
    if (!found && grepl(":", target)) {
      parts <- str_split(target, ":")[[1]]
      reversed <- paste(parts[2], parts[1], sep = ":")
      if (reversed %in% coefs$RawName) {
        row <- coefs[coefs$RawName == reversed, ]
        row$DisplayName <- target
        matched_data[[target]] <- row
        found <- TRUE
      }
    }
    # 策略 3: Compete 模糊匹配
    if (!found && grepl("Compete", target)) {
      target_pattern <- gsub("Compete", "Compete.*", target)
      match_idx <- grep(paste0("^", target_pattern, "$"), coefs$RawName)
      if (length(match_idx) == 0 && grepl(":", target)) {
        parts <- str_split(target, ":")[[1]]
        rev_pattern <- paste0(gsub("Compete", "Compete.*", parts[2]), ":", gsub("Compete", "Compete.*", parts[1]))
        match_idx <- grep(paste0("^", rev_pattern, "$"), coefs$RawName)
      }
      if (length(match_idx) > 0) {
        actual_name <- coefs$RawName[match_idx[1]]
        row <- coefs[coefs$RawName == actual_name, ]
        row$DisplayName <- target
        matched_data[[target]] <- row
        found <- TRUE
      }
    }
  }
  final_df <- do.call(rbind, matched_data)
  if (is.null(final_df)) return(data.frame())
  
  final_df <- final_df %>%
    mutate(
      Variable = factor(DisplayName, levels = rev(target_vars)),
      Estimate = Estimate,
      Lower = Estimate - 1.96 * `Std. Error`,
      Upper = Estimate + 1.96 * `Std. Error`,
      P_val = `Pr(>|t|)`
    )
  return(final_df)
}

# ==============================================================================
# 4. 绘图函数 (新增 R2 计算与展示)
# ==============================================================================
create_forest_plot <- function(model, title_text, target_vars) {
  
  # 1. 计算 R2
  r2_vals <- MuMIn::r.squaredGLMM(model)
  # 提取 Marginal R2 (固定效应解释的部分)
  # r.squaredGLMM 返回一个矩阵: [R2m, R2c]
  r2_m <- round(r2_vals[1, "R2m"], 3)
  r2_c <- round(r2_vals[1, "R2c"], 3) # 如果需要Conditional R2也可以加上
  
  # 构建 subtitle 文本
  subtitle_text <- bquote(italic(R)[m]^2 == .(r2_m))
  
  # 2. 获取数据
  plot_data <- get_model_data(model, target_vars)
  
  if (nrow(plot_data) == 0) return(ggplot() + labs(title = "No Data"))
  
  plot_data$col <- case_when(
    plot_data$P_val < 0.05 ~ "black",
    plot_data$P_val < 0.10 ~ "grey70",
    TRUE                   ~ "white"
  )

  p <- ggplot(plot_data, aes(x = Estimate, y = Variable)) +
    scale_fill_identity() +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey60") +
    geom_errorbarh(aes(xmin = Lower, xmax = Upper), height = 0.2, color = "black") +
    geom_point(size = 2.5, shape = 21, aes(fill = col), color = "black") +
    scale_fill_identity() +
    scale_y_discrete(position = "left") +
    
    # 修改 labs 添加 subtitle
    labs(x = "Standardized Estimate", y = NULL, 
         title = title_text,
         subtitle = subtitle_text) + # 这里添加 subtitle
    
    theme_classic() +
    theme(
      axis.text.y = element_text(size = 10.5, color = "black"),
      plot.title = element_text(hjust = 0.5, size = 11, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5, size = 9, color = "grey30"), # subtitle 样式
      plot.margin = margin(5, 5, 5, 5)
    )
  
  return(p)
}

p1 <- create_forest_plot(lm1, "(a) 1st Year", target_order) + labs(x=NULL)
p2 <- create_forest_plot(lm2, "(b) 2nd Year", target_order)
p3 <- create_forest_plot(lm3, "(c) 3rd Year", target_order) + labs(x=NULL)

p2 <- p2 + theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
p3 <- p3 + theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

final_forest <- p1 + p2 + p3 + plot_layout(ncol = 3, widths = c(1, 1, 1))
print(final_forest)

ggsave("02_output_results_new/figs/figSX_robustness_forest_plot_fixed_compete.pdf", 
       plot = final_forest, width = 8, height = 8)

```

# Relative importance to R2
### Part models----Community-level
```{r}
library(lmerTest)
library(MuMIn)
library(dplyr)
library(ggplot2)
library(tidyr)
library(patchwork)

decompose_triplet_community <- function(dat, t_a, t_b, e_v) {
  default_res <- c(Env=0, Above=0, Below=0, TT=0, TE=0)
  
  f1 <- paste("RGR ~", e_v, "+ (1|sp) + (1|Nenv)")
  f2 <- paste("RGR ~", e_v, "+", t_a, "+ (1|sp) + (1|Nenv)")
  f3 <- paste("RGR ~", e_v, "+", t_a, "+", t_b, "+ (1|sp) + (1|Nenv)")
  f4 <- paste("RGR ~", e_v, "+", t_a, "+", t_b, "+", t_a,":",t_b, "+ (1|sp) + (1|Nenv)")
  f5 <- paste("RGR ~", e_v, "+", t_a, "+", t_b, "+", t_a,":",t_b, "+", t_a,":",e_v, "+", t_b,":",e_v, "+ (1|sp) + (1|Nenv)")
  
  get_r2 <- function(f) {
    mod <- tryCatch({
      lmer(as.formula(f), data = dat, REML = FALSE, control = lmerControl(optimizer = "bobyqa"))
    }, error = function(e) NULL, warning = function(w) NULL)
    
    if (is.null(mod)) return(0)
    
    val <- tryCatch({
      r2_out <- MuMIn::r.squaredGLMM(mod)
      v <- if (is.matrix(r2_out)) r2_out[1, "R2m"] else r2_out["R2m"]
      if (is.na(v) || is.nan(v)) 0 else as.numeric(v)
    }, error = function(e) 0)
    return(val)
  }
  
  r2_vec <- tryCatch({
    c(M1=get_r2(f1), M2=get_r2(f2), M3=get_r2(f3), M4=get_r2(f4), M5=get_r2(f5))
  }, error = function(e) NULL)
  
  if (is.null(r2_vec) || length(r2_vec) != 5 || any(is.na(r2_vec))) return(default_res)
  
  r2_vec <- cummax(r2_vec)
  total_r2 <- r2_vec["M5"]
  
  if (total_r2 < 0.0001) return(default_res)
  
  incr <- c(
    Env   = r2_vec["M1"],
    Above = r2_vec["M2"] - r2_vec["M1"],
    Below = r2_vec["M3"] - r2_vec["M2"],
    TT    = r2_vec["M4"] - r2_vec["M3"],
    TE    = r2_vec["M5"] - r2_vec["M4"]
  )
  return(pmax(0, incr) / total_r2)
}

above_traits <- c("Above_PC1", "Above_PC2", "Above_PC3")
below_traits <- c("Below_PC1", "Below_PC2")
env_vars     <- c("Light", "Moisture", "AP", "Compete")
combinations <- expand.grid(Above = above_traits, Below = below_traits, Env = env_vars, stringsAsFactors = FALSE)

year_datasets <- list("1st" = tda_1st, "2nd" = tda_2nd, "3rd" = tda_3rd)

comm_raw_results <- data.frame()

for (year_name in names(year_datasets)) {
  current_dat <- year_datasets[[year_name]]
  
  for (i in 1:nrow(combinations)) {
    curr_above <- combinations$Above[i]
    curr_below <- combinations$Below[i]
    curr_env   <- combinations$Env[i]
    
    contribs <- decompose_triplet_community(current_dat, curr_above, curr_below, curr_env)
    
    if (is.null(contribs) || length(contribs) != 5 || any(is.na(contribs))) {
      contribs <- c(0, 0, 0, 0, 0)
    }
    
    tmp_raw <- data.frame(
      Year = rep(year_name, 5),
      Combination_ID = rep(i, 5),
      Trait_Above = rep(curr_above, 5),
      Trait_Below = rep(curr_below, 5),
      Env_Var     = rep(curr_env, 5),
      Component = c("Env", "Above", "Below", "TT", "TE"), 
      Value = as.numeric(contribs),
      stringsAsFactors = FALSE
    )
    comm_raw_results <- rbind(comm_raw_results, tmp_raw)
  }
}

library(dplyr)

df_comm <- comm_raw_results %>%
  mutate(
    Level = "Community",
    Component = factor(Component, levels = c("Env","Above","Below","TT","TE")),
    Component_Label = case_when(
      Component == "Env"   ~ "Environment",
      Component == "Above" ~ "Above-ground",
      Component == "Below" ~ "Below-ground",
      Component == "TT"    ~ "Trait*Trait",
      Component == "TE"    ~ "Trait*Env",
      TRUE                 ~ as.character(Component)
    ),
    Component_Label = factor(
      Component_Label,
      levels = c("Environment","Above-ground","Below-ground","Trait*Trait","Trait*Env")
    ),
    Year_Num = case_when(
      Year == "1st" ~ 1,
      Year == "2nd" ~ 2,
      Year == "3rd" ~ 3
    ),
    Combination_ID = factor(Combination_ID)
  ) %>%
  select(Year, Year_Num, Level, Combination_ID, Component, Component_Label, Value)

str(df_comm)
head(df_comm)
library(dplyr)
library(lme4)
library(lmerTest)
library(MuMIn)

df_comm <- comm_raw_results %>%
  mutate(
    Combination_ID = factor(Combination_ID),
    Component = factor(Component, levels = c("Env","Above","Below","TT","TE")),
    Component_Label = case_when(
      Component == "Env"   ~ "Environment",
      Component == "Above" ~ "Above-ground",
      Component == "Below" ~ "Below-ground",
      Component == "TT"    ~ "Trait*Trait",
      Component == "TE"    ~ "Trait*Env",
      TRUE ~ as.character(Component)
    ),
    Component_Label = factor(Component_Label,
      levels = c("Environment","Above-ground","Below-ground","Trait*Trait","Trait*Env")
    ),
    Year_Num = case_when(
      Year == "1st" ~ 1,
      Year == "2nd" ~ 2,
      Year == "3rd" ~ 3
    )
  ) %>%
  select(Year, Year_Num, Combination_ID, Component, Component_Label, Value)

fit_comm_component <- function(dat) {
  m <- lmer(Value ~ Year_Num + (1 | Combination_ID), data = dat, REML = TRUE)
  co <- summary(m)$coefficients["Year_Num", ]
  r2 <- r.squaredGLMM(m)[1]  # marginal R2

  p <- unname(co["Pr(>|t|)"])

  tibble(
    slope = unname(co["Estimate"]),
    se    = unname(co["Std. Error"]),
    p     = p,
    star  = case_when(
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      p < 0.1   ~ ".",
      TRUE      ~ ""
    ),
    R2 = r2
  )
}

res_comm_component <- df_comm %>%
  group_by(Component, Component_Label) %>%
  group_modify(~ fit_comm_component(.x)) %>%
  ungroup()

res_comm_component


library(ggplot2)
library(dplyr)

library(dplyr)

# 如果你表里已经有 star，就直接用；否则按 p 生成
slopes_comm <- res_comm_component %>%
  mutate(
    star = ifelse(!is.na(star), star,
                  case_when(
                    p < 0.001 ~ "***",
                    p < 0.01  ~ "**",
                    p < 0.05  ~ "*",
                    p < 0.1   ~ ".",
                    TRUE      ~ ""
                  )),
    lab = paste0("b=", formatC(slope, format = "f", digits = 3), star),
    x_pos = 3.15
  )


ggplot(df_comm, aes(x = Year_Num, y = Value, color = Component_Label, fill = Component_Label)) +
  geom_smooth(aes(group = Component_Label), method = "lm", se = TRUE, linewidth = 1, alpha = 0.3) +
  scale_color_manual(values = my_colors) +
  scale_fill_manual(values = my_colors) +
  scale_x_continuous(breaks = 1:3, labels = c("1st", "2nd", "3rd")) +
  labs(title = "(c) Community-level", y = NULL, x = "Year") +
  theme_Publication() 












library(lme4)
library(lmerTest)
library(dplyr)
library(tidyr)
library(ggplot2)
library(broom.mixed)
library(patchwork)




# 1.1 Main Effect (Env)
df_env <- comm_raw_results %>% 
  filter(Component == "Env") %>%
  mutate(Specific_Var = Env_Var, Analysis_Group = "Env_Main")

# 1.2 Main Effect (Traits)
df_trait <- comm_raw_results %>% 
  filter(Component %in% c("Above", "Below")) %>%
  mutate(
    Specific_Var = ifelse(Component=="Above", Trait_Above, Trait_Below),
    Analysis_Group = "Trait_Main"
  )

# 1.3 TT Interaction (Split to Traits)
df_tt <- comm_raw_results %>%
  filter(Component == "TT") %>%
  pivot_longer(cols = c(Trait_Above, Trait_Below), values_to = "Specific_Var") %>%
  mutate(Value = Value / 2, Analysis_Group = "TT_Interaction")

# 1.4 TE Interaction (Split to Traits)
df_te_trait <- comm_raw_results %>%
  filter(Component == "TE") %>%
  pivot_longer(cols = c(Trait_Above, Trait_Below), values_to = "Specific_Var") %>%
  mutate(Value = Value / 2, Analysis_Group = "TE_Trait") # 假设均分给性状端

# 1.5 TE Interaction (Split to Env) -> 这里是之前漏掉的关键！
# TE 的贡献完全归因于环境因子 (用于分析环境在交互中的重要性)
# 注意：这里 Value 不除以 2，因为从环境角度看，这是该环境因子参与的所有交互的总和
# 或者为了与性状端量级一致，也可以 /2，取决于“相对重要性”的定义。
# 为了画图量级一致，建议保持一致的处理 (比如都 /2)，或者都用原始值。
# 之前的图中 TE_Env 量级和 TE_Trait 差不多，说明可能都除以了2。
# 我们这里统一：都除以 2，代表“交互作用的一半归因于一方”。
df_te_env <- comm_raw_results %>%
  filter(Component == "TE") %>%
  mutate(
    Specific_Var = Env_Var, 
    Value = Value / 2, 
    Analysis_Group = "TE_Env"
  )

# 合并所有数据用于分析
df_all_analysis <- bind_rows(
  df_env %>% select(Year, Combination_ID, Value, Specific_Var, Analysis_Group),
  df_trait %>% select(Year, Combination_ID, Value, Specific_Var, Analysis_Group),
  df_tt %>% select(Year, Combination_ID, Value, Specific_Var, Analysis_Group),
  df_te_trait %>% select(Year, Combination_ID, Value, Specific_Var, Analysis_Group),
  df_te_env %>% select(Year, Combination_ID, Value, Specific_Var, Analysis_Group)
) %>%
  mutate(
    Year_Num = case_when(Year == "1st" ~ 1, Year == "2nd" ~ 2, Year == "3rd" ~ 3),
    Combination_ID = as.factor(Combination_ID),
    Value = pmax(0, Value) # 防御负值
  )

# ==============================================================================
# 2. 统计检验 (LMM) - 覆盖所有 5 个分析组
# ==============================================================================
run_trend_test_complete <- function(data) {
  unique_groups <- unique(data$Analysis_Group)
  results_list <- list()
  
  for (grp in unique_groups) {
    sub_data <- data %>% filter(Analysis_Group == grp)
    unique_vars <- unique(sub_data$Specific_Var)
    
    for (var in unique_vars) {
      var_data <- sub_data %>% filter(Specific_Var == var)
      
      if (var(var_data$Value, na.rm=TRUE) > 1e-10) {
        mod <- tryCatch({
          lmer(Value ~ Year_Num + (1 | Combination_ID), data = var_data)
        }, error = function(e) NULL)
        
        sig_mark <- ""
        if (!is.null(mod)) {
          res <- broom.mixed::tidy(mod, effects = "fixed") %>% filter(term == "Year_Num")
          if (nrow(res) > 0 && !is.na(res$p.value)) {
            p <- res$p.value
            if (p < 0.001) sig_mark <- "***"
            else if (p < 0.01) sig_mark <- "**"
            else if (p < 0.05) sig_mark <- "*"
            else if (p < 0.1) sig_mark <- "."
          }
        }
        
        results_list[[paste(grp, var, sep="_")]] <- data.frame(
          Analysis_Group = grp, Specific_Var = var, Label = sig_mark, stringsAsFactors = FALSE
        )
      }
    }
  }
  return(do.call(rbind, results_list))
}

print("Running Complete LMM Tests...")
sig_results_complete <- run_trend_test_complete(df_all_analysis)
print(sig_results_complete %>% filter(Label != ""))

# ==============================================================================
# 3. 绘图数据准备
# ==============================================================================
plot_data_agg <- df_all_analysis %>%
  group_by(Year, Analysis_Group, Specific_Var) %>%
  summarise(
    Mean = mean(Value, na.rm=TRUE),
    SE = sd(Value, na.rm=TRUE)/sqrt(n()),
    .groups = "drop"
  )



### 测试不同的变量的时间格局
library(dplyr)
library(lme4)
library(lmerTest)
library(broom.mixed)
library(purrr)

# 1) 确保变量类型正确
df_all_analysis2 <- df_all_analysis %>%
  mutate(
    Year = factor(Year, levels = c("1st","2nd","3rd")),
    Year_Num = as.numeric(Year_Num),
    Combination_ID = factor(Combination_ID),
    Analysis_Group = factor(Analysis_Group),
    Specific_Var = factor(Specific_Var)
  )

# 2) 对每个 group × var 拟合：Value ~ Year_Num + (1|Combination_ID)
fit_one <- function(dat) {
  # 如果某个子集缺少年份或样本不足，则返回 NA
  if (nrow(dat) < 6 || length(unique(dat$Year_Num)) < 2) {
    return(tibble(
      slope = NA_real_, se = NA_real_, df = NA_real_,
      t = NA_real_, p = NA_real_, ci_low = NA_real_, ci_high = NA_real_
    ))
  }

  m <- lmer(Value ~ Year_Num + (1 | Combination_ID), data = dat, REML = TRUE)

  co <- summary(m)$coefficients
  slope <- co["Year_Num", "Estimate"]
  se    <- co["Year_Num", "Std. Error"]
  df    <- co["Year_Num", "df"]
  tval  <- co["Year_Num", "t value"]
  pval  <- co["Year_Num", "Pr(>|t|)"]

  # Wald 95% CI（快速且足够用于趋势汇总；如需更稳健可用 profile）
  ci <- suppressMessages(confint(m, parm = "Year_Num", method = "Wald"))

  tibble(
    slope = slope, se = se, df = df, t = tval, p = pval,
    ci_low = ci[1], ci_high = ci[2]
  )
}

# 3) 批量运行并输出整洁结果表
library(dplyr)
library(lme4)
library(lmerTest)
library(MuMIn)
library(purrr)

# 1) 确保变量类型正确
df_all_analysis2 <- df_all_analysis %>%
  mutate(
    Year_Num = as.numeric(Year_Num),
    Combination_ID = factor(Combination_ID),
    Analysis_Group = factor(Analysis_Group),
    Specific_Var = factor(Specific_Var)
  )

# 2) 单个子集拟合并提取所需统计量
fit_extract <- function(dat) {

  # 数据不足时跳过
  if (nrow(dat) < 6 || length(unique(dat$Year_Num)) < 2) {
    return(tibble(
      slope = NA_real_,
      se = NA_real_,
      p = NA_real_,
      star = NA_character_,
      R2 = NA_real_
    ))
  }

  m <- lmer(Value ~ Year_Num + (1 | Combination_ID),
            data = dat, REML = TRUE)

  co <- summary(m)$coefficients["Year_Num", ]

  slope <- co["Estimate"]
  se    <- co["Std. Error"]
  pval  <- co["Pr(>|t|)"]

  # 显著性星号
  star <- case_when(
    pval < 0.001 ~ "***",
    pval < 0.01  ~ "**",
    pval < 0.05  ~ "*",
    TRUE ~ ""
  )

  # marginal R2（固定效应解释度）
  R2m <- r.squaredGLMM(m)[1]

  tibble(
    slope = slope,
    se = se,
    p = pval,
    star = star,
    R2 = R2m
  )
}

# 3) 批量运行：group × var
res_trend_minimal <- df_all_analysis2 %>%
  group_by(Analysis_Group, Specific_Var) %>%
  group_modify(~ fit_extract(.x)) %>%
  ungroup() %>%
  arrange(Analysis_Group, p)

res_trend_minimal



## 分组测试时间趋势：
library(dplyr)
library(lme4)
library(lmerTest)

df0 <- df_all_analysis %>%
  mutate(
    Year_Num = as.numeric(Year_Num),
    Combination_ID = factor(Combination_ID),
    Specific_Var = factor(Specific_Var),
    Analysis_Group = factor(Analysis_Group)
  )

fit_common_trend <- function(dat) {
  m <- lmer(Value ~ Year_Num + (1 | Combination_ID) + (1 | Specific_Var),
            data = dat, REML = TRUE)

  co <- summary(m)$coefficients["Year_Num", ]
  out <- data.frame(
    slope = unname(co["Estimate"]),
    se    = unname(co["Std. Error"]),
    p     = unname(co["Pr(>|t|)"])
  )
  out$star <- ifelse(out$p < 0.001, "***",
                     ifelse(out$p < 0.01, "**",
                            ifelse(out$p < 0.05, "*", "")))
  return(list(model = m, stats = out))
}

dat_above <- df0 %>%
  filter(Analysis_Group == "Trait_Main",
         Specific_Var %in% c("Above_PC1", "Above_PC2", "Above_PC3"))

res_above <- fit_common_trend(dat_above)

res_above$stats   # slope / se / p / star
summary(res_above$model)

dat_below <- df0 %>%
  filter(Analysis_Group == "Trait_Main",
         Specific_Var %in% c("Below_PC1", "Below_PC2"))

res_below <- fit_common_trend(dat_below)

res_below$stats
summary(res_below$model)

dat_env3 <- df0 %>%
  filter(Analysis_Group == "Env_Main",
         Specific_Var %in% c("Light", "Moisture", "AP"))

res_env3 <- fit_common_trend(dat_env3)
res_env3$stats
summary(res_env3$model)


dat_tt_above <- df0 %>%
  filter(
    Analysis_Group == "TT_Interaction",
    Specific_Var %in% c(
      "Above_PC1",
      "Above_PC2",
      "Above_PC3",
      "Below_PC1",
      "Below_PC2"
    )
  )
m_tt_above <- lmer(
  Value ~ Year_Num + (1 | Combination_ID) + (1|Specific_Var),
  data = dat_tt_above,
  REML = TRUE
)

summary(m_tt_above)

# 标签准备
label_data <- plot_data_agg %>%
  group_by(Analysis_Group, Specific_Var) %>%
  summarise(Max_Y = max(Mean + SE, na.rm=TRUE), .groups="drop") %>%
  left_join(sig_results_complete, by = c("Analysis_Group", "Specific_Var")) %>%
  mutate(Year = "2nd", Label = ifelse(is.na(Label), "", Label))

# ==============================================================================
# 4. 绘图函数与拼图
# ==============================================================================
plot_group <- function(grp_name, title, remove_x=FALSE) {
  d <- plot_data_agg %>% filter(Analysis_Group == grp_name)
  l <- label_data %>% filter(Analysis_Group == grp_name)
  
  # 强制排序 Env
  if (grepl("Env", grp_name)) {
    levs <- c("Light", "Moisture", "AP", "Compete")
    d$Specific_Var <- factor(d$Specific_Var, levels = levs)
    l$Specific_Var <- factor(l$Specific_Var, levels = levs)
  }
  
  p <- ggplot(d, aes(x = Specific_Var, y = Mean, color = Year, group = Year)) +
    geom_errorbar(aes(ymin = pmax(0, Mean - SE), ymax = Mean + SE), 
                  width = 0.2, position = position_dodge(width = 0.6)) +
    geom_point(position = position_dodge(width = 0.6), size = 3) +
    geom_text(data = l, aes(x = Specific_Var, y = Max_Y * 1.15, label = Label), 
              inherit.aes = FALSE, size = 5, color = "black", fontface = "bold") +
    scale_color_manual(values = c("1st"="#FFE7CC", "2nd"="#F8CBA6", "3rd"="#E08E6D")) +
    labs(title = title, x = NULL, y = "Relative Importance") +
    theme_Publication() +
    theme(legend.position = "none")
  
  if (remove_x) p <- p + theme(axis.text.x = element_blank())
  else p <- p + theme(axis.text.x = element_text(angle=30, hjust=1, color="black"))
  
  return(p)
}

## ============================================================
## Build panels (a–e) with full x-axis labels for traits and envs
## ============================================================

library(ggplot2)

## ---- 1) Full labels ----
trait_x_labels <- c(
  "Above_PC1" = "Leaf economics\n(Above_PC1)",
  "Above_PC2" = "Plant size\n(Above_PC2)",
  "Above_PC3" = "Leaf area\n(Above_PC3)",
  "Below_PC1" = "Root collaboration\n(Below_PC1)",
  "Below_PC2" = "Root economics\n(Below_PC2)"
)

env_x_labels <- c(
  "Light"    = "Light",
  "Moisture" = "Soil moisture",
  "AP"       = "Soil phosphorus",
  "Compete"  = "Competition"
)

## ---- 2) Apply x-axis helper functions ----
apply_trait_x_axis <- function(p, angle = 35, hjust = 1, vjust = 1, size = 9.5) {
  p +
    scale_x_discrete(labels = trait_x_labels) +
    theme(
      axis.text.x = element_text(
        angle = angle, hjust = hjust, vjust = vjust,
        size = size, colour = "black"
      )
    )
}

apply_env_x_axis <- function(p, angle = 0, hjust = 0.5, vjust = 1, size = 9.5) {
  p +
    scale_x_discrete(labels = env_x_labels) +
    theme(
      axis.text.x = element_text(
        angle = angle, hjust = hjust, vjust = vjust,
        size = size, colour = "black"
      )
    )
}

## ---- 3) Common margins (avoid repetition) ----
panel_margin <- theme(plot.margin = margin(t = 1, r = 1, b = 4, l = 1, unit = "pt"))

## ---- 4) Generate panels ----
p_a <- plot_group("Trait_Main", "(a) Trait Main Effect", remove_x = TRUE) +
  ylab(NULL) + panel_margin




p_b <- plot_group("TT_Interaction", "(b) Trait × Trait Interaction", remove_x = TRUE) +
  panel_margin

## (c) x = trait -> full trait labels
p_c <- plot_group("TE_Trait", "(c) Trait × Env Interaction (by Trait)", remove_x = FALSE) +
  ylab(NULL) + panel_margin
p_c <- apply_trait_x_axis(p_c, angle = 35, hjust = 0.8, size = 9.5)

p_d <- plot_group("Env_Main", "(d) Env Main Effect", remove_x = TRUE) +
  ylab(NULL) + panel_margin

## (e) x = env -> full env labels
p_e <- plot_group("TE_Env", "(e) Trait × Env Interaction (by Env)", remove_x = FALSE) +
  ylab(NULL) + panel_margin
p_e <- apply_env_x_axis(p_e, angle = 15, hjust = 0.8, size = 9.5)

# 拼图
fig4a_e <- (p_a / p_b / p_c) | (p_d / p_e + plot_layout(heights = c(1, 1)))
fig4a_e <- fig4a_e + plot_layout(widths = c(1.3, 1), guides = "collect") & theme(legend.position = "bottom")

p_a1 <- plot_group("Trait_Main", "(a) Trait Main Effect", remove_x = TRUE) +
  ylab(NULL) + panel_margin+ylim(c(0,0.11))

library(eoffice)
topptx(p_a1,filename = "02_output_results_new/figs/fig4a1.pptx", width = 2, height = 1)

```

## Species-level
```{r}
library(lme4)
library(lmerTest)
library(MuMIn)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)

# ==============================================================================
# 1. 定义核心分解函数
# ==============================================================================
decompose_triplet <- function(dat, t_a, t_b, e_v) {
  # 简化的模型公式 (注意：物种水平通常不需要 (1|Nenv)，除非有多个环境)
  # 这里保留 (1|Nenv) 以防数据结构确实包含嵌套环境
  f1 <- paste("RGR_scale ~", e_v, "+ (1|Nenv)")
  f2 <- paste("RGR_scale ~", e_v, "+", t_a, "+ (1|Nenv)")
  f3 <- paste("RGR_scale ~", e_v, "+", t_a, "+", t_b, "+ (1|Nenv)")
  f4 <- paste("RGR_scale ~", e_v, "+", t_a, "+", t_b, "+", t_a,":",t_b, "+ (1|Nenv)")
  f5 <- paste("RGR_scale ~", e_v, "+", t_a, "+", t_b, "+", t_a,":",t_b, "+", t_a,":",e_v, "+", t_b,":",e_v, "+ (1|Nenv)")
  
  get_r2 <- function(f) {
    mod <- tryCatch({
      lmer(as.formula(f), data = dat, REML = FALSE, control = lmerControl(optimizer = "bobyqa"))
    }, error = function(e) NULL, warning = function(w) NULL)
    
    if (is.null(mod)) return(0)
    
    val <- tryCatch({
      r2_out <- MuMIn::r.squaredGLMM(mod)
      if (is.matrix(r2_out)) r2_out[1, "R2m"] else r2_out["R2m"]
    }, error = function(e) 0)
    return(if (is.na(val)) 0 else as.numeric(val))
  }
  
  r2_vec <- c(M1=get_r2(f1), M2=get_r2(f2), M3=get_r2(f3), M4=get_r2(f4), M5=get_r2(f5))
  r2_vec[is.na(r2_vec)] <- 0
  r2_vec <- cummax(r2_vec) # 确保递增
  
  total_r2 <- r2_vec["M5"]
  if(is.na(total_r2) || total_r2 < 0.0001) return(rep(0, 5)) 
  
  incr <- c(
    Env   = r2_vec["M1"],
    Above = r2_vec["M2"] - r2_vec["M1"],
    Below = r2_vec["M3"] - r2_vec["M2"],
    TT    = r2_vec["M4"] - r2_vec["M3"],
    TE    = r2_vec["M5"] - r2_vec["M4"]
  )
  return(pmax(0, incr)) # 确保非负
}

# ==============================================================================
# 2. 准备循环变量
# ==============================================================================
# 确保使用 RGR_scale (标准化后的生长速率)
above_traits <- c("Above_PC1", "Above_PC2", "Above_PC3")
below_traits <- c("Below_PC1", "Below_PC2")
env_vars     <- c("Light", "Moisture", "AP", "Compete") # 注意 Compete 列名
combinations <- expand.grid(Above = above_traits, Below = below_traits, Env = env_vars, stringsAsFactors = FALSE)

year_datasets <- list("1st" = tda_1st, "2nd" = tda_2nd, "3rd" = tda_3rd)
species_list <- unique(tda_1st$sp)

results_list <- list()
counter <- 1

for (year_name in names(year_datasets)) {
  current_dat <- year_datasets[[year_name]]
  
  for (sp_name in species_list) {
    sp_dat <- current_dat[current_dat$sp == sp_name, ]
  
    if (nrow(sp_dat) < 15) next 
    for (i in 1:nrow(combinations)) {
      raw_incr <- decompose_triplet(sp_dat, combinations$Above[i], combinations$Below[i], combinations$Env[i])
      total_r2 <- sum(raw_incr, na.rm = TRUE)
      
      if(total_r2 > 0.001) {

        rel_contrib <- raw_incr / total_r2
        results_list[[counter]] <- data.frame(
          Year = year_name,
          Species = sp_name,
          Combination_ID = i,
          Trait_Above = combinations$Above[i],
          Trait_Below = combinations$Below[i],
          Env_Var = combinations$Env[i],
          Val_Env   = as.numeric(rel_contrib[1]),
          Val_Above = as.numeric(rel_contrib[2]),
          Val_Below = as.numeric(rel_contrib[3]),
          Val_TT    = as.numeric(rel_contrib[4]),
          Val_TE    = as.numeric(rel_contrib[5]),
          Total_R2 = total_r2,
          stringsAsFactors = FALSE
        )
        counter <- counter + 1
      }
    }
  }
}
print("Loop Completed.")
species_raw_detailed <- bind_rows(results_list)

df_community <- comm_raw_results %>%
  mutate(
    Species = "Community",
    Level = "Community") %>%
  select(Year, Level, Species, Combination_ID, Component, Value)%>%
  mutate(
    Component_Label = case_when(
      Component == "Env"   ~ "Environment",
      Component == "Above" ~ "Above-ground",
      Component == "Below" ~ "Below-ground",
      Component == "TT"    ~ "Trait*Trait",
      Component == "TE"    ~ "Trait*Env",
      TRUE                 ~ Component
    ),
    Component_Label = factor(Component_Label, 
       levels = c("Environment", "Above-ground", "Below-ground", "Trait*Trait", "Trait*Env")),
    # 年份数值化
    Year_Num = case_when(
      Year == "1st" ~ 1,
      Year == "2nd" ~ 2,
      Year == "3rd" ~ 3
    )
  )

df_species <- species_raw_detailed %>%
  select(Year, Species, Combination_ID, Val_Env, Val_Above, Val_Below, Val_TT, Val_TE) %>%
  pivot_longer(
    cols = starts_with("Val_"),
    names_to = "Component",
    values_to = "Value"
  ) %>%
  mutate(
    Level = "Species",
    Component = recode(Component,
      "Val_Env"   = "Env",
      "Val_Above" = "Above",
      "Val_Below" = "Below",
      "Val_TT"    = "TT",
      "Val_TE"    = "TE"
    )) %>%
  select(Year,Level, Species, Combination_ID, Component, Value)%>%
  mutate(
    # 映射全称标签
    Component_Label = case_when(
      Component == "Env"   ~ "Environment",
      Component == "Above" ~ "Above-ground",
      Component == "Below" ~ "Below-ground",
      Component == "TT"    ~ "Trait*Trait",
      Component == "TE"    ~ "Trait*Env",
      TRUE                 ~ Component
    ),
    # 设置因子顺序
    Component_Label = factor(Component_Label, 
       levels = c("Environment", "Above-ground", "Below-ground", "Trait*Trait", "Trait*Env")),
    # 年份数值化
    Year_Num = case_when(
      Year == "1st" ~ 1,
      Year == "2nd" ~ 2,
      Year == "3rd" ~ 3
    )
  )


library(dplyr)
library(lme4)
library(lmerTest)
library(MuMIn)

df_species2 <- df_species %>%
  mutate(
    Species = factor(Species),
    Combination_ID = factor(Combination_ID),
    Component = factor(Component)
  )

fit_component_trend <- function(dat) {

  m <- lmer(
    Value ~ Year_Num + (1 | Species) + (1 | Combination_ID),
    data = dat,
    REML = TRUE
  )

  co <- summary(m)$coefficients["Year_Num", ]

  r2 <- r.squaredGLMM(m)[1]  # marginal R2

  tibble(
    slope = unname(co["Estimate"]),
    se    = unname(co["Std. Error"]),
    p     = unname(co["Pr(>|t|)"]),
    star  = case_when(
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      p < 0.1   ~ ".",
      TRUE      ~ ""
    ),
    R2 = r2
  )
}

res_species_component <- df_species2 %>%
  group_by(Component, Component_Label) %>%
  group_modify(~ fit_component_trend(.x)) %>%
  ungroup()
res_species_component


my_colors <- c(
  "Above-ground" = "#A5B68D",
  "Below-ground" = "#AB886D",
  "Trait*Trait"  = "#FFE1AF",
  "Environment"  = "#7C93C3",
  "Trait*Env"    = "#C2D9FF"
)


level_order <- c("Above-ground", "Below-ground", "Trait*Trait", "Environment", "Trait*Env")
df_species$Component_Label <- factor(df_species$Component_Label, levels = level_order)
df_community$Component_Label <- factor(df_community$Component_Label, levels = level_order)

my_colors <- c(
  "Environment"  = "#7C93C3",
  "Above-ground" = "#A5B68D",
  "Below-ground" = "#AB886D",
  "Trait*Trait"  = "#FFE1AF",
  "Trait*Env"    = "#C2D9FF")
p_ac <- ggplot(df_community, aes(x = Year_Num, y = Value, color = Component_Label, fill = Component_Label)) +
  geom_jitter(width = 0.25, alpha = 0.2, size = 0.8, shape = 16) +
  geom_smooth(aes(group=Component_Label),method = "lm", se = TRUE, size = 1, alpha = 0.3) +
  scale_color_manual(values = my_colors) +
  scale_fill_manual(values = my_colors) +
  scale_x_continuous(breaks = 1:3, labels = c("1st", "2nd", "3rd")) +
  labs(title = "(b) Temporal Trends", y = "Relative Importance", x = NULL) +
  theme_classic(base_size = 13) + 
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 11)
  ) #+facet_wrap(~Component, nrow = 1, scales = "fixed")
p_ac

library(dplyr)
library(broom)

library(dplyr)
library(broom)

slopes_species <- df_species %>%
  mutate(Year_Num = as.numeric(Year_Num)) %>%
  group_by(Component_Label) %>%
  do({
    m <- lm(Value ~ Year_Num, data = .)
    broom::tidy(m, conf.int = TRUE) %>% 
      filter(term == "Year_Num")
  }) %>%
  ungroup() %>%
  transmute(
    Component_Label,
    slope = estimate,
    conf.low,
    conf.high,
    p.value)

comp_levels <- levels(df_species$Component_Label)

slopes_species <- slopes_species %>%
  mutate(
    Component_Label = factor(Component_Label, levels = comp_levels),
    lab = sprintf("slope = %.3f", slope)   # 只显示 slope
   
  )

y_top <- mean(df_species$Value, na.rm = TRUE)
y_rng <- diff(range(df_species$Value, na.rm = TRUE))

slopes_species <- slopes_species %>%
  arrange(Component_Label) %>%
  mutate(
    x_pos = 2.95,  
    y_pos = y_top - (row_number() - 1) * 0.03 * y_rng)

library(ggplot2)

p_bs <- ggplot(df_species, aes(x = Year_Num, y = Value, color = Component_Label, fill = Component_Label)) +
  geom_smooth(aes(group = Component_Label), method = "lm", se = TRUE, linewidth = 1, alpha = 0.3) +
  scale_color_manual(values = my_colors) +
  scale_fill_manual(values = my_colors) +
  scale_x_continuous(breaks = 1:3, labels = c("1st", "2nd", "3rd")) +
  labs(title = "(d) Species-level", y = NULL, x = "Year") +
  theme_Publication() + panel_margin +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 11)
  ) +
  # 标注 slope：文字颜色跟线一致
  geom_text(
    data = slopes_species,
    aes(x = x_pos, y = y_pos, label = lab, color = Component_Label),
    inherit.aes = FALSE,
    hjust = 1,        # 右对齐，贴近右侧
    size = 3.2
  ) +
  # 给右侧一点空间避免文字被裁剪
  coord_cartesian(xlim = c(1, 3.2))

p_bs
p_bs1 <- ggplot(df_species, aes(x = Year_Num, y = Value, color = Component_Label, fill = Component_Label)) +
  #geom_jitter(width = 0.25, alpha = 0.2, size = 0.8, shape = 16) +
  geom_smooth(aes(group=Component_Label),method = "lm", se = TRUE, size = 1, alpha = 0.3) +
  scale_color_manual(values = my_colors) +
  scale_fill_manual(values = my_colors) +
  scale_x_continuous(breaks = 1:3, labels = c("1st", "2nd", "3rd")) +
  labs(title = "(d) Species-level", y = NULL, x = "Year") +
  theme_Publication() + panel_margin + 
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 11)
  ) 

p_bs
p_ac + p_bs +plot_layout(ncol=2)

fig4a_f <- p_a + p_d + p_b + p_e + p_c + p_bs + plot_layout(ncol=2)
fig4a_f <- fig4a_f + plot_layout(widths = c(1, 1), guides = "collect") & theme(legend.position = "bottom")

print(fig4a_f )
ggsave("02_output_results_new/figs/fig4_sp_var_imp_final.pdf", plot = fig4a_f, width = 7, height = 8)

library(eoffice)
topptx(fig4a_f,filename = "02_output_results_new/figs/fig4_sp_var_imp_final.pptx", width = 7, height = 8)


# Figure 4
fig4_a_panel <- p_a      # 图4(a): Trait Main Effect
fig4_b_panel <- p_b      # 图4(b): Trait × Trait Interaction

```

# Results based on full model

## Figure 5
```{r}
library(lmerTest)
library(sjPlot)
total_scaled_tda1 <- pcatraits_tda1
lme_full <- lmerTest::lmer(
  RGR ~ Above_PC1 + Above_PC2 + Above_PC3 + Below_PC1 + Below_PC2 +
    Light + Moisture + AP + Compete + growth_time +
    Above_PC1:growth_time + Above_PC2:growth_time + Above_PC3:growth_time + 
    Below_PC1:growth_time + Below_PC2:growth_time +
    Light:growth_time + Moisture:growth_time + AP:growth_time + Compete:growth_time +
    Above_PC1:Below_PC1 +  Above_PC2:Below_PC1 +  Above_PC3:Below_PC1 +
    Above_PC1:Below_PC2 +  Above_PC2:Below_PC2 +  Above_PC3:Below_PC2 +
    Below_PC1:Below_PC2 + Above_PC1:Above_PC2 + Above_PC1:Above_PC3 + Above_PC2:Above_PC3 +
    Above_PC1:Light + Above_PC1:Moisture + Above_PC1:AP + Above_PC1:Compete +
    Above_PC2:Light + Above_PC2:Moisture + Above_PC2:AP + Above_PC2:Compete +
    Above_PC3:Light + Above_PC3:Moisture + Above_PC3:AP + Above_PC3:Compete +
    Below_PC1:Light + Below_PC1:Moisture + Below_PC1:AP + Below_PC1:Compete +
    Below_PC2:Light + Below_PC2:Moisture + Below_PC2:AP + Below_PC2:Compete +
    Above_PC1:Below_PC1:growth_time +  Above_PC2:Below_PC1:growth_time +  Above_PC3:Below_PC1:growth_time +
    Above_PC1:Below_PC2:growth_time +  Above_PC2:Below_PC2:growth_time +  Above_PC3:Below_PC2:growth_time +
    Below_PC1:Below_PC2:growth_time + Above_PC1:Above_PC2:growth_time + Above_PC1:Above_PC3:growth_time +
    Above_PC2:Above_PC3:growth_time +
    Above_PC1:Light:growth_time + Above_PC1:Moisture:growth_time + Above_PC1:AP:growth_time +
    Above_PC1:Compete:growth_time +
    Above_PC2:Light:growth_time + Above_PC2:Moisture:growth_time + Above_PC2:AP:growth_time +
    Above_PC2:Compete:growth_time +
    Above_PC3:Light:growth_time + Above_PC3:Moisture:growth_time + Above_PC3:AP:growth_time +
    Above_PC3:Compete:growth_time +
    Below_PC1:Light:growth_time + Below_PC1:Moisture:growth_time + Below_PC1:AP:growth_time +
    Below_PC1:Compete:growth_time +
    Below_PC2:Light:growth_time + Below_PC2:Moisture:growth_time + Below_PC2:AP:growth_time +
    Below_PC2:Compete:growth_time +
    (1 | sp) + (1 | Nenv),
  data = total_scaled_tda1
)
tab_model(lme_full)
## ============================================================
## 1) Forest plot: robust term-finding (handles factor like Compete)
## ============================================================

forest_var_paired_sig_offset2 <- function(model,
                                          vars,
                                          var_labels,
                                          time_var = "growth_time",
                                          title = "(a)",
                                          xlab = "Effect size",
                                          y_offset = 0.15) {

  cf <- fixef(model)
  cn <- names(cf)
  ci <- confint(model, method = "Wald")
  sm <- summary(model)$coefficients

  # helper: swap A:B <-> B:A
  swap_term <- function(x) {
    if (!grepl(":", x)) return(x)
    paste(rev(strsplit(x, ":")[[1]]), collapse = ":")
  }

  # helper: find exact or swapped; if not found, allow prefix match for factor terms
  find_term_robust <- function(var, type = c("Main", "Time")) {
    type <- match.arg(type)

    # target interaction term string (ideal)
    term0 <- if (type == "Main") var else paste0(var, ":", time_var)

    # 1) exact
    if (term0 %in% cn) return(term0)
    # 2) swapped
    term1 <- swap_term(term0)
    if (term1 %in% cn) return(term1)

    # 3) prefix match (for factor-coded variables, e.g., Competeinter)
    # Main: grep("^Compete", cn) but exclude ":" interactions
    # Time: grep("^Compete.*:growth_time$" or "^growth_time:Compete", etc.)
    if (type == "Main") {
      hits <- grep(paste0("^", var), cn, value = TRUE)
      hits <- hits[!grepl(":", hits)]
      # remove intercept if someone passes "(Intercept)" by mistake
      hits <- hits[hits != "(Intercept)"]
      if (length(hits) == 1) return(hits)
      if (length(hits) > 1) {
        message("Multiple main-effect terms matched for ", var, ": ",
                paste(hits, collapse = ", "),
                ". Using the first: ", hits[1])
        return(hits[1])
      }
    } else {
      # allow both var:*:growth_time and growth_time:var*
      hits <- grep(paste0("^", var, ".*:", time_var, "$"), cn, value = TRUE)
      if (length(hits) == 0) {
        hits <- grep(paste0("^", time_var, ":", var), cn, value = TRUE)
      }
      if (length(hits) == 1) return(hits)
      if (length(hits) > 1) {
        message("Multiple time-interaction terms matched for ", var, ": ",
                paste(hits, collapse = ", "),
                ". Using the first: ", hits[1])
        return(hits[1])
      }
    }

    stop("Term not found for variable '", var, "' (", type, ").\n",
         "Available coefficient names include:\n",
         paste(head(cn, 40), collapse = ", "),
         if (length(cn) > 40) "\n... (truncated)" else "")
  }

  # build y-axis table
  var_df <- tibble(
    Var = vars,
    label = var_labels,
    y_base = rev(seq_along(vars))
  )

  # extract one row per (Var, Effect)
  get_effect <- function(v, type) {
    term <- find_term_robust(v, type)
    tibble(
      Var = v,
      Effect = type,
      term = term,
      estimate = unname(cf[term]),
      lower = ci[term, 1],
      upper = ci[term, 2],
      p = sm[term, "Pr(>|t|)"]
    )
  }

  df <- bind_rows(
    lapply(vars, get_effect, type = "Main"),
    lapply(vars, get_effect, type = "Time")
  ) %>%
    left_join(var_df, by = "Var") %>%
    mutate(
      y = ifelse(Effect == "Main", y_base + y_offset, y_base - y_offset),
      Sig = case_when(
        p < 0.05 ~ "p < 0.05",
        p < 0.10 ~ "0.05–0.10",
        TRUE     ~ "p ≥ 0.10"
      ),
      Effect = factor(Effect, levels = c("Main", "Time"))
    )

  ggplot(df, aes(x = estimate, y = y)) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_errorbarh(aes(xmin = lower, xmax = upper, colour = Effect),
                   height = 0.12, linewidth = 0.6) +
    geom_point(aes(fill = Sig, colour = Effect),
               shape = 21, size = 2.2, stroke = 0.9) +
    scale_fill_manual(
      values = c("p < 0.05" = "black", "0.05–0.10" = "grey70", "p ≥ 0.10" = "white"),
      name = "P value"
    ) +
    scale_color_manual(values = c("Main" = "black", "Time" = "#F3A26D"),
                       labels = c("Main", "Var × Year"),
                       name = NULL) +
    scale_y_continuous(breaks = var_df$y_base, labels = var_df$label) +
    labs(title = title, x = xlab, y = NULL) +
    theme_Publication() + 
  theme(
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background  = element_rect(fill = "white", colour = NA),

    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
}

vars9 <- c("Above_PC1","Above_PC2","Above_PC3","Below_PC1","Below_PC2",
           "Light","Moisture","AP","Compete")

labels9 <- c("Leaf economics\n(Above_PC1)",
             "Plant size\n(Above_PC2)",
             "Leaf area\n(Above_PC3)",
             "Root collaboration\n(Below_PC1)",
             "Root economics\n(Below_PC2)",
             "Light intensity",
             "Soil Moisture",
             "Soil phosphorus",
             "Competition")

forest9 <- forest_var_paired_sig_offset2(
  model = lme_full,
  vars = vars9,
  var_labels = labels9,
  title = "(a)"
)

forest9

```

```{r,fig.height=9,fig.width=9}
library(ggeffects)
year_cols <- c("#F5E7C6","#F7B980","#DE802B")
plot_trait_year_raw <- function(model_raw,
                                data,
                                trait,
                                time_var = "growth_time",
                                response_raw = "RGR",
                                time_vals,
                                time_labels = c("1st", "2nd", "3rd"),
                                trait_range = NULL,
                                title = NULL,
                                colors = year_cols,
                                show_ci = TRUE) {

  pred <- ggpredict(
    model_raw,
    terms = c(
      trait,
      sprintf("%s [%s]", time_var, paste(time_vals, collapse = ", "))
    ),
    type = "fixed"
  )

  pred_df <- as.data.frame(pred) %>%
    mutate(
      Year = factor(group,
                    levels = as.character(time_vals),
                    labels = time_labels)
    )

  pal <- setNames(colors, time_labels)

  p <- ggplot() +
    geom_point(
      data = data,
      aes(x = .data[[trait]], y = .data[[response_raw]],colour = comp_year),
      alpha  = 0.1,
      size   = 0.2
    )+
    scale_color_manual(values = colors)

  if (show_ci) {
    p <- p + geom_ribbon(
      data = pred_df,
      aes(x = x, ymin = conf.low, ymax = conf.high, fill = Year),
      alpha = 0.15,
      colour = NA
    )
  }

  p <- p +
    geom_line(
      data = pred_df,
      aes(x = x, y = predicted, colour = Year),
      linewidth = 1.2
    ) +
    scale_color_manual(values = pal, name = "Year") +
    scale_fill_manual(values  = pal, name = "Year") +
    labs(x = trait, y = "RGR", title = title) +
    theme_Publication(base_size=11)+
    theme(
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10))+
  scale_y_continuous(
    labels = number_format(accuracy = 0.1)
  ) +
  scale_x_continuous(
    breaks = function(x) floor(min(x)):ceiling(max(x))
  )

  if (!is.null(trait_range)) {
    p <- p + coord_cartesian(xlim = trait_range, ylim = c(0, 1.2))
  } else {
    p <- p + coord_cartesian(ylim = c(0, NA))
  }

  return(p)
}

gt_vals <- sort(as.numeric(unique(total_scaled_tda1$growth_time)))
gt_vals

lme_plot_raw=lme_full

trait_vars <- c(
  "Above_PC1",
  "Above_PC2",
  "Above_PC3",
  "Below_PC1",
  "Below_PC2")

trait_ranges <- lapply(
  total_scaled_tda1[trait_vars],
  function(x) quantile(x, probs = c(0.001,0.999), na.rm = TRUE))

trait_ranges

above_pc1_p <- plot_trait_year_raw(
  model_raw   = lme_plot_raw,
  data        = total_scaled_tda1,
  trait       = "Above_PC1",
  time_vals   = gt_vals,
  trait_range = trait_ranges$Above_PC1,
  title       = "(b)") + xlab("Leaf economics")+ylab(NULL)+
  scale_x_continuous(breaks = c(-4,-2, 0, 2,4) )+
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "pt"))


above_pc2_p <- plot_trait_year_raw(lme_plot_raw, total_scaled_tda1, "Above_PC2",
                                  time_vals = gt_vals,
                                  trait_range = trait_ranges$Above_PC2,
                                  title="(c)")+ylab(NULL) + xlab("Plant size") +
  scale_x_continuous(breaks = c(-4,-2, 0, 2,4) )+
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "pt"))


above_pc3_p <- plot_trait_year_raw(lme_plot_raw, total_scaled_tda1, "Above_PC3",
                                  time_vals = gt_vals,
                                  trait_range = trait_ranges$Above_PC3,
                                  title="(d)")+ylab(NULL)+ xlab("Leaf area")+
  scale_x_continuous(breaks = c(-4,-2, 0, 2,4) )+
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "pt"))


below_pc1_p <- plot_trait_year_raw(lme_plot_raw, total_scaled_tda1, "Below_PC1",
                                  time_vals = gt_vals,
                                  trait_range = trait_ranges$Below_PC1,
                                  title="(e)") + xlab("Root collaboration") +
  scale_x_continuous(breaks = c(-4,-2, 0, 2,4) ) +
  ylab(expression("Relative growth rate (RGR, log cm yr"^{-1}*")"))+
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "pt"))


below_pc2_p <- plot_trait_year_raw(lme_plot_raw, total_scaled_tda1, "Below_PC2",
                                  time_vals = gt_vals,
                                  trait_range = trait_ranges$Below_PC2,
                                  title="(f)")+ylab(NULL)+ xlab("Root economics") +
  scale_x_continuous(breaks = c(-4,-2, 0, 2,4) )+
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "pt"))


plot_env_year <- function(model,
                          data,
                          env,
                          response = "RGR",
                          time_var = "growth_time",
                          time_vals,
                          time_labels = c("1st", "2nd", "3rd"),
                          env_range = NULL,         
                          title = NULL,
                          colors = year_cols,
                          show_points = TRUE,
                          point_alpha = 0.20,
                          point_size  = 0.6,
                          ci_alpha = 0.12) {

  pred <- ggpredict(
    model,
    terms = c(
      env,
      sprintf("%s [%s]", time_var, paste(time_vals, collapse = ", "))
    ),
    type = "fixed"
  )

  pred_df <- as.data.frame(pred) %>%
    mutate(
      Year = factor(group,
                    levels = as.character(time_vals),
                    labels = time_labels)
    )

  pal <- setNames(colors, time_labels)

  p <- ggplot()
  if (show_points) {
    p <- p + 
    geom_point(
      data = data,
      aes(x = .data[[env]], y = .data[[response]],colour = comp_year),
      alpha  = 0.1,
      size   = 0.2
    )+scale_color_manual(values = colors)

  }
  p <- p +
    geom_ribbon(
      data = pred_df,
      aes(x = x, ymin = conf.low, ymax = conf.high, fill = Year),
      alpha = ci_alpha,
      colour = NA
    ) +
    geom_line(
      data = pred_df,
      aes(x = x, y = predicted, colour = Year),
      linewidth = 1.2
    ) +
    scale_color_manual(values = pal, name = "Year") +
    scale_fill_manual(values  = pal, name = "Year") +
    labs(x = env, y = "RGR", title = title) +
    theme_Publication(base_size = 11) +
    theme(
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10))+
  scale_y_continuous(
    labels = number_format(accuracy = 0.1)) 
  # 坐标裁剪：env 范围 + y>=0（如果你用原始RGR，通常应>=0）
  if (!is.null(env_range)) {
    p <- p + coord_cartesian(xlim = env_range, ylim = c(0, NA))
  } else {
    p <- p + coord_cartesian(ylim = c(0, NA))
  }

  return(p)
}

lme_env_raw=lme_full
env_ranges <- list(
  Light    = quantile(total_scaled_tda1$Light,    c(0.05, 0.95), na.rm = TRUE),
  Moisture = quantile(total_scaled_tda1$Moisture, c(0.05, 0.95), na.rm = TRUE),
  AP       = quantile(total_scaled_tda1$AP,       c(0.05, 0.95), na.rm = TRUE)
)
p_light <- plot_env_year(
  model = lme_env_raw,
  data  = total_scaled_tda1,
  env   = "Light",
  time_vals = gt_vals,
  env_range = env_ranges$Light,
  title = "(g)")+ylab(NULL)+scale_x_continuous(breaks = c(-1, 0, 1) )+theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "pt"))+xlab("Light intensity")


p_moist <- plot_env_year(
  lme_env_raw, total_scaled_tda1,
  env = "Moisture",
  time_vals = gt_vals,
  env_range = env_ranges$Moisture,
  title = "(h)"
)+theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "pt"))+ylab(NULL)+xlab("Soil moisture")

p_ap <- plot_env_year(
  lme_env_raw, total_scaled_tda1,
  env = "AP",
  time_vals = gt_vals,
  env_range = env_ranges$AP,
  title = "(i)"
)+ylab(NULL)+theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "pt"))+xlab("Soil phosphorus")

p_compete <- plot_env_year(
  lme_env_raw, total_scaled_tda1,
  env = "Compete",
  time_vals = gt_vals,
  title = "(j)",
  show_points = TRUE
)+ylab(NULL)+theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "pt"))+xlab("Competition")

fig5a <- forest9 + theme_Publication(base_size = 11) +
    theme(
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 11))

## ---- Left column: (a) only uses 70% height ----
fig5a_short <- (
  fig5a +
    theme(
      legend.position = "none" 
    )
) 

## ---- Right block: 3×3 panels (unchanged) ----
p_right <- (above_pc1_p + above_pc2_p + above_pc3_p +
              below_pc1_p + below_pc2_p + p_light +
              p_moist + p_ap + p_compete) +
  plot_layout(ncol = 3) &
  theme(legend.position = "bottom")

## ---- Combine left + right with width ratio 1.2 : 3 ----
fig5_all <- (fig5a_short | p_right) +
  plot_layout(widths = c(1, 3), guides = "collect") &
  theme(
    legend.position = "bottom",
    legend.text  = element_text(size = 10),
    legend.title = element_text(size = 10)
  )

fig5_all

ggsave(fig5_all,
       file = "02_output_results_new/figs/fig5_trait_effect.pdf",
       height = 6.1, width = 8.1, bg = "white")

topptx(fig5_all,
       filename = "02_output_results_new/figs/fig5_trait_effect.pptx",
       height = 8, width = 10.5)
```

## Figure 6 Trait * Env * Year
```{r,eval=FALSE}
TE_terms <- c(
  "Above_PC1:Moisture",
  "Above_PC1:AP",
  "Above_PC2:Light",
  "Above_PC2:Moisture",
  "Above_PC2:Competeinter",
  "Above_PC3:Light",
  "Above_PC3:AP",
  "Above_PC3:Competeinter",
  "Below_PC1:Light",
  "Below_PC1:Competeinter",
  "Below_PC2:Light",
  "Below_PC2:Moisture"
)
pair_labels_TE <- c(
  "Above_PC1:Moisture"     = "Leaf econ × Moisture",
  "Above_PC1:AP"           = "Leaf econ × AP",
  "Above_PC2:Light"        = "Plant size × Light",
  "Above_PC2:Moisture"     = "Plant size × Moisture",
  "Above_PC2:Competeinter" = "Plant size × Competition",
  "Above_PC3:Light"        = "Leaf area × Light",
  "Above_PC3:AP"           = "Leaf area × AP",
  "Above_PC3:Competeinter" = "Leaf area × Competition",
  "Below_PC1:Light"        = "Root collab × Light",
  "Below_PC1:Competeinter" = "Root collab × Competition",
  "Below_PC2:Light"        = "Root econ × Light",
  "Below_PC2:Moisture"     = "Root econ × Moisture"
)
tab_model(lme_full)
## Figure 6 Trait × Env × Year  (Forest plot)
library(dplyr)
library(ggplot2)

forest_TE_paired_sig3 <- function(model,
                                  TE_terms,
                                  pair_labels,
                                  time_var = "growth_time",
                                  title = "Trait–environment interactions and their temporal change",
                                  xlab  = "Effect size (estimate ± 95% CI)",
                                  y_offset = 0.18,
                                  col_TE = "black",
                                  col_TETwitter = "#F3A26D") {

  ## ---------- helpers ----------
  swap_term <- function(x) {
    if (!grepl(":", x)) return(x)
    paste(rev(strsplit(x, ":")[[1]]), collapse = ":")
  }

  find_term <- function(term, names_vec) {
    if (term %in% names_vec) return(term)
    alt <- swap_term(term)
    if (alt %in% names_vec) return(alt)
    stop("Term not found: ", term)
  }

  sig_class <- function(p) {
    dplyr::case_when(
      p < 0.05 ~ "p < 0.05",
      p < 0.10 ~ "0.05–0.10",
      TRUE     ~ "p ≥ 0.10"
    )
  }

  ## ---------- extract ----------
  cf <- fixef(model)
  ci <- confint(model, method = "Wald")
  sm <- summary(model)$coefficients
  pvals <- sm[, "Pr(>|t|)"]

  y_map <- tibble(
    TE = TE_terms,
    label = pair_labels[TE_terms],
    y_base = rev(seq_along(TE_terms))
  )

  get_row <- function(te, is_time = FALSE) {
    term <- if (!is_time) te else paste0(te, ":", time_var)
    term <- find_term(term, names(cf))
    p <- unname(pvals[term])

    tibble(
      TE = te,
      EffectType = ifelse(is_time, "TE × Year", "TE"),
      estimate = unname(cf[term]),
      lower = ci[term, 1],
      upper = ci[term, 2],
      p = p,
      Sig = sig_class(p)
    )
  }

  df <- bind_rows(
    lapply(TE_terms, get_row, is_time = FALSE),
    lapply(TE_terms, get_row, is_time = TRUE)
  ) %>%
    left_join(y_map, by = "TE") %>%
    mutate(
      y = ifelse(EffectType == "TE",
                 y_base + y_offset,
                 y_base - y_offset),
      ## edge color encodes effect type only
      Edge = ifelse(EffectType == "TE", col_TE, col_TETwitter)
    )

  ## ---------- plot ----------
  ggplot(df, aes(x = estimate, y = y)) +
    geom_vline(xintercept = 0, linetype = "dashed") +

    geom_errorbarh(
      aes(xmin = lower, xmax = upper, colour = EffectType),
      height = 0.15,
      linewidth = 0.6
    ) +

    geom_point(
      aes(fill = Sig, colour = EffectType),
      shape = 21,
      size = 2,
      stroke = 0.9
    ) +

    ## fill = significance (same as Fig.5)
    scale_fill_manual(
      values = c(
        "p < 0.05"  = "black",
        "0.05–0.10" = "grey70",
        "p ≥ 0.10"  = "white"
      ),
      name = "P value"
    ) +

    ## outline = effect type (black vs orange)
    scale_colour_manual(
      values = c("TE" = col_TE, "TE × Year" = col_TETwitter),
      name = NULL,
      labels = c("Trait × environment", "Trait × environment × year")
    ) +

    scale_y_continuous(
      breaks = y_map$y_base,
      labels = y_map$label
    ) +

    labs(title = title, x = xlab, y = NULL) +
    theme_Publication()
}

## ---- run ----
res_TE <- forest_TE_paired_sig3(
  model = lme_full,
  TE_terms = TE_terms,
  pair_labels = pair_labels_TE,
  time_var = "growth_time",
  title = "Trait–environment interactions"
)

res_TE


```

## Figure 6 Trait * Trait * Year
```{r,eval=FALSE}
library(dplyr)
library(ggplot2)

forest_TT_paired_sig3 <- function(model,
                                  TT_terms,
                                  pair_labels,
                                  time_var = "growth_time",
                                  title = "Trait–trait interactions and their temporal change",
                                  xlab  = "Effect size (estimate ± 95% CI)",
                                  y_offset = 0.18,
                                  col_TT = "black",
                                  col_TTYear = "#F3A26D") {

  ## ---------- helpers ----------
  swap_term <- function(x) {
    if (!grepl(":", x)) return(x)
    paste(rev(strsplit(x, ":")[[1]]), collapse = ":")
  }

  find_term <- function(term, names_vec) {
    if (term %in% names_vec) return(term)
    alt <- swap_term(term)
    if (alt %in% names_vec) return(alt)
    stop("Term not found: ", term)
  }

  sig_class <- function(p) {
    dplyr::case_when(
      p < 0.05 ~ "p < 0.05",
      p < 0.10 ~ "0.05–0.10",
      TRUE     ~ "p ≥ 0.10"
    )
  }

  ## ---------- extract ----------
  cf <- fixef(model)
  ci <- confint(model, method = "Wald")
  sm <- summary(model)$coefficients
  pvals <- sm[, "Pr(>|t|)"]

  y_map <- tibble(
    TT = TT_terms,
    label = pair_labels[TT_terms],
    y_base = rev(seq_along(TT_terms))
  )

  get_row <- function(tt, is_time = FALSE) {
    term <- if (!is_time) tt else paste0(tt, ":", time_var)
    term <- find_term(term, names(cf))
    p <- unname(pvals[term])

    tibble(
      TT = tt,
      EffectType = ifelse(is_time, "TT × Year", "TT"),
      estimate = unname(cf[term]),
      lower = ci[term, 1],
      upper = ci[term, 2],
      p = p,
      Sig = sig_class(p)
    )
  }

  df <- bind_rows(
    lapply(TT_terms, get_row, is_time = FALSE),
    lapply(TT_terms, get_row, is_time = TRUE)
  ) %>%
    left_join(y_map, by = "TT") %>%
    mutate(
      y = ifelse(EffectType == "TT",
                 y_base + y_offset,
                 y_base - y_offset),
      EffectType = factor(EffectType, levels = c("TT", "TT × Year"))
    )

  ## ---------- plot ----------
  ggplot(df, aes(x = estimate, y = y)) +
    geom_vline(xintercept = 0, linetype = "dashed") +

    geom_errorbarh(
      aes(xmin = lower, xmax = upper, colour = EffectType),
      height = 0.12,
      linewidth = 0.8
    ) +

    geom_point(
      aes(fill = Sig, colour = EffectType),
      shape = 21,
      size = 2,
      stroke = 0.9
    ) +

    ## fill = significance (same as Fig.5)
    scale_fill_manual(
      values = c(
        "p < 0.05"  = "black",
        "0.05–0.10" = "grey70",
        "p ≥ 0.10"  = "white"
      ),
      name = "P value"
    ) +

    ## outline/errorbar = effect type (black vs orange)
    scale_colour_manual(
      values = c("TT" = col_TT, "TT × Year" = col_TTYear),
      name = NULL,
      labels = c("Trait × trait", "Trait × trait × year")
    ) +

    scale_y_continuous(
      breaks = y_map$y_base,
      labels = y_map$label
    ) +

    labs(title = title, x = xlab, y = NULL) +
    theme_Publication()
}

## ===========================
## 1) Define TT terms (unchanged)
## ===========================
TT_terms <- c(
  "Above_PC1:Above_PC2",
  "Above_PC1:Above_PC3",
  "Above_PC1:Below_PC1",
  "Above_PC1:Below_PC2",
  "Above_PC2:Above_PC3",
  "Above_PC2:Below_PC1",
  "Above_PC2:Below_PC2",
  "Above_PC3:Below_PC1",
  "Above_PC3:Below_PC2",
  "Below_PC1:Below_PC2"
)

pair_labels <- c(
  "Above_PC1:Above_PC2" = "Leaf econ × Plant size",
  "Above_PC1:Above_PC3" = "Leaf econ × Leaf area",
  "Above_PC1:Below_PC1" = "Leaf econ × Root collab",
  "Above_PC1:Below_PC2" = "Leaf econ × Root econ",
  "Above_PC2:Above_PC3" = "Plant size × Leaf area",
  "Above_PC2:Below_PC1" = "Plant size × Root collab",
  "Above_PC2:Below_PC2" = "Plant size × Root econ",
  "Above_PC3:Below_PC1" = "Leaf area × Root collab",
  "Above_PC3:Below_PC2" = "Leaf area × Root econ",
  "Below_PC1:Below_PC2" = "Root collab × Root econ"
)

## ===========================
## 2) Run & plot
## ===========================
res_TT <- forest_TT_paired_sig3(
  model = lme_full,
  TT_terms = TT_terms,
  pair_labels = pair_labels,
  time_var = "growth_time",
  title = "Trait–trait interactions"
)

res_TT

res_TT+res_TE+plot_layout(ncol=2)
```

## Figure 6 TE/TT_selected
```{r}
library(dplyr)
library(ggplot2)

forest_TE_paired_sig_filter <- function(model,
                                        TE_terms,
                                        pair_labels,
                                        time_var = "growth_time",
                                        title = "Trait–environment interactions",
                                        xlab  = "Effect size",
                                        y_offset = 0.12,
                                        col_TE = "black",
                                        col_TEYear = "#F3A26D",
                                        p_keep = 0.10) {

  # --- helpers ---
  swap_term <- function(x) {
    if (!grepl(":", x)) return(x)
    paste(rev(strsplit(x, ":")[[1]]), collapse = ":")
  }
  find_term <- function(term, names_vec) {
    if (term %in% names_vec) return(term)
    alt <- swap_term(term)
    if (alt %in% names_vec) return(alt)
    stop("Term not found: ", term)
  }
  sig_class <- function(p) {
    case_when(
      p < 0.05 ~ "p < 0.05",
      p < 0.10 ~ "0.05–0.10",
      TRUE     ~ "p ≥ 0.10"
    )
  }

  # --- extract ---
  cf <- fixef(model)
  ci <- confint(model, method = "Wald")
  sm <- summary(model)$coefficients
  pvals <- sm[, "Pr(>|t|)"]

  get_row <- function(te, is_time = FALSE) {
    term <- if (!is_time) te else paste0(te, ":", time_var)
    term <- find_term(term, names(cf))
    p <- unname(pvals[term])

    tibble(
      Pair = te,
      EffectType = ifelse(is_time, "TE × Year", "TE"),
      estimate = unname(cf[term]),
      lower = ci[term, 1],
      upper = ci[term, 2],
      p = p,
      Sig = sig_class(p)
    )
  }

  df <- bind_rows(
    lapply(TE_terms, get_row, is_time = FALSE),
    lapply(TE_terms, get_row, is_time = TRUE)
  )

  # --- filter: keep pair if at least one of (TE, TE×Year) has p < p_keep ---
  keep_pairs <- df %>%
    group_by(Pair) %>%
    summarise(min_p = min(p, na.rm = TRUE), .groups = "drop") %>%
    filter(min_p < p_keep) %>%
    pull(Pair)

  df <- df %>% filter(Pair %in% keep_pairs)

  # if nothing left, fail gracefully
  if (nrow(df) == 0) {
    stop("No TE pairs passed the filter (min p < ", p_keep, ").")
  }

  # --- rebuild y map for filtered pairs (preserve original TE_terms order) ---
  keep_pairs <- TE_terms[TE_terms %in% keep_pairs]
  y_map <- tibble(
    Pair = keep_pairs,
    label = pair_labels[keep_pairs],
    y_base = rev(seq_along(keep_pairs))
  )

  df <- df %>%
    left_join(y_map, by = "Pair") %>%
    mutate(
      y = ifelse(EffectType == "TE", y_base + y_offset, y_base - y_offset),
      EffectType = factor(EffectType, levels = c("TE", "TE × Year"))
    )

  ggplot(df, aes(x = estimate, y = y)) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_errorbarh(
      aes(xmin = lower, xmax = upper, colour = EffectType),
      height = 0.15, linewidth = 0.6
    ) +
    geom_point(
      aes(fill = Sig, colour = EffectType),
      shape = 21, size = 2, stroke = 0.9
    ) +
    scale_fill_manual(
      values = c(
        "p < 0.05"  = "black",
        "0.05–0.10" = "grey70",
        "p ≥ 0.10"  = "white"
      ),
      name = "P value"
    ) +
    scale_colour_manual(
      values = c("TE" = col_TE, "TE × Year" = col_TEYear),
      name = NULL,
      labels = c("Trait × environment", "Trait × environment × year")
    ) +
    scale_y_continuous(breaks = y_map$y_base, labels = y_map$label) +
    labs(title = title, x = xlab, y = NULL) +
    theme_Publication()
}


forest_TT_paired_sig_filter <- function(model,
                                        TT_terms,
                                        pair_labels,
                                        time_var = "growth_time",
                                        title = "Trait–trait interactions",
                                        xlab  = "Effect size",
                                        y_offset = 0.12,
                                        col_TT = "black",
                                        col_TTYear = "#F3A26D",
                                        p_keep = 0.10) {

  swap_term <- function(x) {
    if (!grepl(":", x)) return(x)
    paste(rev(strsplit(x, ":")[[1]]), collapse = ":")
  }
  find_term <- function(term, names_vec) {
    if (term %in% names_vec) return(term)
    alt <- swap_term(term)
    if (alt %in% names_vec) return(alt)
    stop("Term not found: ", term)
  }
  sig_class <- function(p) {
    case_when(
      p < 0.05 ~ "p < 0.05",
      p < 0.10 ~ "0.05–0.10",
      TRUE     ~ "p ≥ 0.10"
    )
  }

  cf <- fixef(model)
  ci <- confint(model, method = "Wald")
  sm <- summary(model)$coefficients
  pvals <- sm[, "Pr(>|t|)"]

  get_row <- function(tt, is_time = FALSE) {
    term <- if (!is_time) tt else paste0(tt, ":", time_var)
    term <- find_term(term, names(cf))
    p <- unname(pvals[term])

    tibble(
      Pair = tt,
      EffectType = ifelse(is_time, "TT × Year", "TT"),
      estimate = unname(cf[term]),
      lower = ci[term, 1],
      upper = ci[term, 2],
      p = p,
      Sig = sig_class(p)
    )
  }

  df <- bind_rows(
    lapply(TT_terms, get_row, is_time = FALSE),
    lapply(TT_terms, get_row, is_time = TRUE)
  )

  # --- filter pairs ---
  keep_pairs <- df %>%
    group_by(Pair) %>%
    summarise(min_p = min(p, na.rm = TRUE), .groups = "drop") %>%
    filter(min_p < p_keep) %>%
    pull(Pair)

  df <- df %>% filter(Pair %in% keep_pairs)

  if (nrow(df) == 0) {
    stop("No TT pairs passed the filter (min p < ", p_keep, ").")
  }

  # preserve original order
  keep_pairs <- TT_terms[TT_terms %in% keep_pairs]
  y_map <- tibble(
    Pair = keep_pairs,
    label = pair_labels[keep_pairs],
    y_base = rev(seq_along(keep_pairs))
  )

  df <- df %>%
    left_join(y_map, by = "Pair") %>%
    mutate(
      y = ifelse(EffectType == "TT", y_base + y_offset, y_base - y_offset),
      EffectType = factor(EffectType, levels = c("TT", "TT × Year"))
    )

  ggplot(df, aes(x = estimate, y = y)) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_errorbarh(
      aes(xmin = lower, xmax = upper, colour = EffectType),
      height = 0.12, linewidth = 0.8
    ) +
    geom_point(
      aes(fill = Sig, colour = EffectType),
      shape = 21, size = 2, stroke = 0.9
    ) +
    scale_fill_manual(
      values = c(
        "p < 0.05"  = "black",
        "0.05–0.10" = "grey70",
        "p ≥ 0.10"  = "white"
      ),
      name = "P value"
    ) +
    scale_colour_manual(
      values = c("TT" = col_TT, "TT × Year" = col_TTYear),
      name = NULL,
      labels = c("Trait × trait", "Trait × trait × year")
    ) +
    scale_y_continuous(breaks = y_map$y_base, labels = y_map$label) +
    labs(title = title, x = xlab, y = NULL) +
    theme_Publication()
}


plot_3way_time2 <- function(model,
                            data,
                            x,                    
                            mod,                  
                            time = "growth_time",
                            time_vals = c(-1.7059989,-0.2379998,1.2299992),
                            time_labels = c("1st year", "2nd year", "3rd year"),
                            mod_vals_num = c(-1, 0, 1),
                            mod_labels_num = c("Low", "Medium", "High"),
                            mod_vals_fac = NULL,           # 默认自动取 levels(data[[mod]])
                            mod_labels_fac = NULL,         # 默认用 levels
                            x_label = NULL,
                            y_label = "Predicted RGR",
                            title = NULL,
                            palette = c("#B7B2D0", "#e3a561", "#448870"),
                            ribbon_alpha = 0.18) {

  suppressPackageStartupMessages({
    library(ggeffects)
    library(dplyr)
    library(ggplot2)
  })

  stopifnot(x %in% names(data), mod %in% names(data), time %in% names(data))

  if (is.null(x_label)) x_label <- x
  if (is.null(title)) title <- sprintf("(%s × %s) × %s", x, mod, time)

  mod_is_factor <- is.factor(data[[mod]]) || is.character(data[[mod]])

  # --- 构造 ggpredict 的 terms ---
  if (!mod_is_factor) {
    # 连续：用数值水平
    term_mod <- sprintf("%s [%s]", mod, paste(mod_vals_num, collapse = ", "))
    group_levels <- as.character(mod_vals_num)
    group_labels <- mod_labels_num
    pal <- palette
  } else {
    # 因子：用水平名
    if (is.null(mod_vals_fac)) mod_vals_fac <- levels(as.factor(data[[mod]]))
    if (is.null(mod_labels_fac)) mod_labels_fac <- mod_vals_fac

    term_mod <- sprintf("%s [%s]", mod, paste(mod_vals_fac, collapse = ", "))
    group_levels <- mod_vals_fac
    group_labels <- mod_labels_fac

    # 因子只有两类时，取 palette 前两色
    pal <- palette[seq_len(length(mod_vals_fac))]
  }

  term_time <- sprintf("%s [%s]", time, paste(time_vals, collapse = ", "))

  
  x_range <- as.numeric(quantile(data[[x]], probs = c(0.05, 0.95), na.rm = TRUE))
  
  pred <- ggpredict(model,
                    terms = c(x, term_mod, term_time),
                    type = "fixed")

  df <- as.data.frame(pred) %>%
    mutate(
      Mod = factor(group, levels = group_levels, labels = group_labels),
      Year = factor(facet, levels = as.character(time_vals), labels = time_labels)
    )

  ggplot(df, aes(x = x, y = predicted, colour = Mod, fill = Mod)) +
    geom_line(linewidth = 1.1) +
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
                alpha = ribbon_alpha, colour = NA) +
    facet_wrap(~Year, nrow = 1) +
    scale_color_manual(values = pal) +
    scale_fill_manual(values = pal) +
    labs(x = x_label, y = y_label, colour = mod, fill = mod, title = title) +
    coord_cartesian(xlim = x_range)+
    theme_Publication()
}

plot_3way_time2 <- function(model,
                            data,
                            x,
                            mod,
                            time = "growth_time",
                            time_vals = c(-1.7059989,-0.2379998,1.2299992),
                            time_labels = c("1st year", "2nd year", "3rd year"),

                            mod_vals_num = c(-1, 0, 1),
                            mod_labels_num = c("Low", "Medium", "High"),

                            mod_vals_fac = NULL,
                            mod_labels_fac = NULL,

                            x_label = NULL,
                            y_label = "Predicted RGR",
                            title = NULL,
                            palette = c("#B7B2D0", "#e3a561", "#448870"),
                            ribbon_alpha = 0.18,

                            # ✅新增：散点设置
                            response_raw = "RGR",
                            show_points = TRUE,
                            point_alpha = 0.10,
                            point_size  = 1,
                            points_by_mod = FALSE,     # TRUE=点按Mod分组上色；FALSE=点全灰
                            time_round_digits = 6      # ✅关键：解决浮点匹配问题
                            ) {

  suppressPackageStartupMessages({
    library(ggeffects)
    library(dplyr)
    library(ggplot2)
  })

  stopifnot(x %in% names(data), mod %in% names(data), time %in% names(data))
  stopifnot(response_raw %in% names(data))

  if (is.null(x_label)) x_label <- x
  if (is.null(title)) title <- sprintf("(%s × %s) × %s", x, mod, time)

  mod_is_factor <- is.factor(data[[mod]]) || is.character(data[[mod]])

  # --- 构造 ggpredict 的 terms ---
  if (!mod_is_factor) {
    term_mod <- sprintf("%s [%s]", mod, paste(mod_vals_num, collapse = ", "))
    group_levels <- as.character(mod_vals_num)
    group_labels <- mod_labels_num
    pal <- palette
  } else {
    if (is.null(mod_vals_fac)) mod_vals_fac <- levels(as.factor(data[[mod]]))
    if (is.null(mod_labels_fac)) mod_labels_fac <- mod_vals_fac

    term_mod <- sprintf("%s [%s]", mod, paste(mod_vals_fac, collapse = ", "))
    group_levels <- mod_vals_fac
    group_labels <- mod_labels_fac
    pal <- palette[seq_len(length(mod_vals_fac))]
  }

  term_time <- sprintf("%s [%s]", time, paste(time_vals, collapse = ", "))

  # x 轴用 5-95%
  x_range <- as.numeric(quantile(data[[x]], probs = c(0.05, 0.95), na.rm = TRUE))

  # ---- 预测 ----
  pred <- ggpredict(model,
                    terms = c(x, term_mod, term_time),
                    type = "fixed")

  df <- as.data.frame(pred) %>%
    mutate(
      Mod  = factor(group, levels = group_levels, labels = group_labels),
      Year = factor(as.character(facet),
                    levels = as.character(time_vals),
                    labels = time_labels)
    )

  # =========================================================
  # ✅ 给原始点贴 Year（按年分面）——用 round 防止浮点匹配失败
  # =========================================================
  year_map <- tibble(
    gt = round(as.numeric(time_vals), time_round_digits),
    Year = factor(gt, levels = round(as.numeric(time_vals), time_round_digits),
                  labels = time_labels)
  )

  data_pt <- data %>%
    mutate(gt = round(as.numeric(.data[[time]]), time_round_digits)) %>%
    left_join(year_map, by = "gt") %>%
    filter(!is.na(Year))

  # （可选）点也按 Mod 分组着色（连续 mod 用三分位；因子 mod 用水平）
  if (show_points && points_by_mod && !mod_is_factor) {
    cuts <- quantile(data_pt[[mod]], probs = c(1/3, 2/3), na.rm = TRUE)
    data_pt <- data_pt %>%
      mutate(
        Mod_pt = case_when(
          .data[[mod]] <= cuts[1] ~ mod_labels_num[1],
          .data[[mod]] <= cuts[2] ~ mod_labels_num[2],
          TRUE ~ mod_labels_num[3]
        ),
        Mod_pt = factor(Mod_pt, levels = mod_labels_num)
      )
  }
  if (show_points && points_by_mod && mod_is_factor) {
    data_pt <- data_pt %>%
      mutate(Mod_pt = factor(.data[[mod]], levels = group_levels, labels = group_labels))
  }

  # =========================================================
  # ✅ 开始画图：先点，再CI，再线（点不会盖住线）
  # =========================================================
  p <- ggplot()

  if (show_points) {
    if (points_by_mod) {
      p <- p + geom_point(
        data = data_pt,
        aes(x = .data[[x]], y = .data[[response_raw]], colour = Mod_pt),
        alpha = point_alpha,
        size  = point_size)
    } else {
      p <- p +
        geom_point(
        data = data_pt,
        aes(x = .data[[x]], y = .data[[response_raw]]),
        colour = "grey60",
        alpha  = point_alpha,
        size   = point_size
      )
    }
  }

  p <- p +
    # geom_ribbon(
    #   data = df,
    #   aes(x = x, ymin = conf.low, ymax = conf.high, fill = Mod),
    #   alpha = ribbon_alpha, colour = NA
    # ) +
    geom_line(
      data = df,
      aes(x = x, y = predicted, colour = Mod),
      linewidth = 1.1
    ) +
    facet_wrap(~Year, nrow = 1) +
    scale_color_manual(values = pal) +
    scale_fill_manual(values = pal) +
    labs(x = x_label, y = y_label, colour = mod, fill = mod, title = title) +
    coord_cartesian(xlim = x_range) +
    theme_Publication() +
  theme(
    strip.background = element_rect(fill = "white", colour = NA),
    strip.text       = element_text(colour = "black")
  )

  return(p)
}


p_tt1 <- plot_3way_time2(
  model = lme_full,
  data  = total_scaled_tda1,
  x = "Above_PC1",
  mod = "Below_PC2",
  mod_vals_num = c(-1.5, 0, 1.5),
  mod_labels_num = c("low", "Medium", "high plant size"),
  x_label = "Leaf economics (Above_PC1)",
  title = "(c) Leaf economics × Root economics")+ylim(c(0,0.7))+theme(legend.position = "none")+
  theme(plot.margin = margin(t = 1, r = 1, b = 4, l = 1, unit = "pt"))
p_tt1

p_bpc2_m <- plot_3way_time2(
  model = lme_full,
  data  = total_scaled_tda1,
  x = "Below_PC2",
  mod = "Moisture",
  mod_vals_num = c(-1.5, 0, 1.5),
  mod_labels_num = c("Low", "Medium", "High light"),
  x_label = "Root economics (Below_PC2)",
  title = "(d) Root economics × Soil moisture"
)+ylim(c(0,0.75))+theme(legend.position = "none")+
  theme(plot.margin = margin(t = 1, r = 1, b = 4, l = 1, unit = "pt"))
p_bpc2_m

p_bpc1_l <- plot_3way_time2(
  model = lme_full,
  data  = total_scaled_tda1,
  x = "Below_PC1",
  mod = "Light",
  mod_vals_num = c(-1.5, 0, 1.5),
  mod_labels_num = c("Low", "Medium", "High light"),
  x_label = "Root collaboration (Below_PC1)",
  title = "(e) Root collaboration × Light"
)+ylim(c(0,0.7))+theme(legend.position = "none")+
  theme(plot.margin = margin(t = 1, r = 1, b = 4, l = 1, unit = "pt"))


res_TE_f <- forest_TE_paired_sig_filter(
  model = lme_full,
  TE_terms = TE_terms,
  pair_labels = pair_labels_TE,
  time_var = "growth_time",
  p_keep = 0.10,
  title = "(b) Trait–environment interactions"
)+theme(axis.text.y = element_text(size = 10)) +
  theme(plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "pt"))

res_TT_f <- forest_TT_paired_sig_filter(
  model = lme_full,
  TT_terms = TT_terms,
  pair_labels = pair_labels,
  time_var = "growth_time",
  p_keep = 0.10,
  title = "(a) Trait–trait interactions"
)+ xlab(NULL)+theme(axis.text.y = element_text(size = 10)) + 
  theme(plot.margin = margin(t = 1, r = 1, b = 4, l = 1, unit = "pt"))

## ---- Left column: (a) over (b) ----
p_left <- (res_TT_f / res_TE_f) +
  plot_layout(
    heights = c(4, 10),
    guides = "collect"
  )

## ---- Right column: three panels stacked ----
p_right <- (p_tt1 + p_bpc1_l + p_bpc2_m) +
  plot_layout(ncol = 1)

## ---- Combine left and right (1 : 2 width) ----
fig6_all <- (p_left | p_right) +
  plot_layout(
    widths = c(0.7, 2),
    guides = "collect"
  ) &
  theme(
    legend.position = "bottom"
  )

fig6_all

ggsave(
  filename = "02_output_results_new/figs/fig6_combined.pdf",
  plot     = fig6_all,
  width    = 9,      # 建议宽度（英寸）
  height   = 7,       # 建议高度（英寸）
  units    = "in",
  bg       = "white"  # 强制白底，避免透明
)
```

# New Figure 4

```{r}
library(dplyr)
library(ggplot2)
library(patchwork)
library(ggeffects)
library(grid)
library(scales)
library(eoffice)

## ============================================================
## 0) 统一的一些标签与样式
## ============================================================

panel_margin <- theme(
  plot.margin = margin(t = 3, r = 3, b = 5, l = 3, unit = "pt")
)

trait_x_labels <- c(
  "Above_PC1" = "Leaf economics\n(Above_PC1)",
  "Above_PC2" = "Plant size\n(Above_PC2)",
  "Above_PC3" = "Leaf area\n(Above_PC3)",
  "Below_PC1" = "Root collaboration\n(Below_PC1)",
  "Below_PC2" = "Root economics\n(Below_PC2)"
)

apply_trait_x_axis <- function(p, angle = 10, hjust = 1, vjust = 1, size = 9.5) {
  p +
    scale_x_discrete(labels = trait_x_labels) +
    theme(
      axis.text.x = element_text(
        angle = angle, hjust = hjust, vjust = vjust,
        size = size, colour = "black"
      )
    )
}

## ============================================================
## 1) 左侧两幅图：从原始绘图函数开始重建
##    依赖对象：plot_data_agg, label_data
## ============================================================

plot_group <- function(grp_name, title, remove_x = FALSE) {
  d <- plot_data_agg %>% filter(Analysis_Group == grp_name)
  l <- label_data %>% filter(Analysis_Group == grp_name)

  p <- ggplot(d, aes(x = Specific_Var, y = Mean, color = Year, group = Year)) +
    geom_errorbar(
      aes(ymin = pmax(0, Mean - SE), ymax = Mean + SE),
      width = 0.2, position = position_dodge(width = 0.7)
    ) +
    geom_point(position = position_dodge(width = 0.7), size = 4) +
    geom_text(
      data = l,
      aes(x = Specific_Var, y = Max_Y * 1.15, label = Label),
      inherit.aes = FALSE,
      size = 5,
      color = "black",
      fontface = "bold"
    ) +
    scale_color_manual(values = c("1st" = "#FFE7CC", "2nd" = "#F8CBA6", "3rd" = "#E08E6D")) +
    labs(title = title, x = NULL, y = "Relative Importance") +
    theme_Publication() +
    theme(legend.position = "none")

  if (remove_x) {
    p <- p + theme(axis.text.x = element_blank())
  } else {
    p <- p + theme(axis.text.x = element_text(angle = 10, hjust = 1, color = "black"))
  }

  p
}

## 左侧(a)
fig4_a_panel <- plot_group("Trait_Main", "(a) Trait Main Effect", remove_x = TRUE) +
  ylab(NULL) +
  panel_margin

## 左侧(b) —— 按你的要求添加 x 轴标题
fig4_b_panel <- plot_group("TT_Interaction", "(b) Trait × Trait Interaction", remove_x = FALSE) +
  labs(x = "Trait dimensions", y = "Relative Importance") +
  panel_margin

fig4_b_panel <- apply_trait_x_axis(fig4_b_panel, angle = 20, hjust = 1, size = 9.3)

## ============================================================
## 2) 右侧上半部分：图5(b)(c)(d)(e)(f) 从原始函数开始重建
##    依赖对象：lme_full, total_scaled_tda1
## ============================================================

year_cols <- c("#F5E7C6", "#F7B980", "#DE802B")

plot_trait_year_raw <- function(model_raw,
                                data,
                                trait,
                                time_var = "growth_time",
                                response_raw = "RGR",
                                time_vals,
                                time_labels = c("1st", "2nd", "3rd"),
                                trait_range = NULL,
                                title = NULL,
                                colors = year_cols,
                                show_ci = TRUE) {

  pred <- ggpredict(
    model_raw,
    terms = c(
      trait,
      sprintf("%s [%s]", time_var, paste(time_vals, collapse = ", "))
    ),
    type = "fixed"
  )

  pred_df <- as.data.frame(pred) %>%
    mutate(
      Year = factor(group,
                    levels = as.character(time_vals),
                    labels = time_labels)
    )

  pal <- setNames(colors, time_labels)

  p <- ggplot() +
    # geom_point(
    #   data = data,
    #   aes(x = .data[[trait]], y = .data[[response_raw]], colour = comp_year),
    #   alpha = 0.10,
    #   size  = 0.05
    # ) +
    scale_color_manual(values = colors)

  if (show_ci) {
    p <- p + geom_ribbon(
      data = pred_df,
      aes(x = x, ymin = conf.low, ymax = conf.high, fill = Year),
      alpha = 0.15,
      colour = NA
    )
  }

  p <- p +
    geom_line(
      data = pred_df,
      aes(x = x, y = predicted, colour = Year),
      linewidth = 1.2
    ) +
    scale_color_manual(values = pal, name = "Year") +
    scale_fill_manual(values = pal, name = "Year") +
    labs(x = trait, y = "RGR", title = title) +
    theme_Publication(base_size = 11) +
    theme(
      axis.text.x = element_text(size = 10),
      axis.text.y = element_text(size = 10),
      legend.position = "none"
    ) +
    scale_y_continuous(labels = number_format(accuracy = 0.1)) +
    scale_x_continuous(
      breaks = function(x) floor(min(x)):ceiling(max(x))
    )

  if (!is.null(trait_range)) {
    p <- p + coord_cartesian(xlim = trait_range, ylim = c(0, 1.2))
  } else {
    p <- p + coord_cartesian(ylim = c(0, 1.2))
  }

  p
}

gt_vals <- sort(as.numeric(unique(total_scaled_tda1$growth_time)))

trait_vars <- c("Above_PC1", "Above_PC2", "Above_PC3", "Below_PC1", "Below_PC2")
trait_ranges <- lapply(
  total_scaled_tda1[trait_vars],
  function(x) quantile(x, probs = c(0.001, 0.999), na.rm = TRUE)
)

## 图5(b)
fig5_b_panel <- plot_trait_year_raw(
  model_raw   = lme_full,
  data        = total_scaled_tda1,
  trait       = "Above_PC1",
  time_vals   = gt_vals,
  trait_range = trait_ranges$Above_PC1,
  title       = "(c)"
) +
  xlab("Leaf economics") + ylab(NULL) +
  scale_x_continuous(breaks = c(-4, -2, 0, 2, 4)) +
  panel_margin

## 图5(c)
fig5_c_panel <- plot_trait_year_raw(
  model_raw   = lme_full,
  data        = total_scaled_tda1,
  trait       = "Above_PC2",
  time_vals   = gt_vals,
  trait_range = trait_ranges$Above_PC2,
  title       = "(d)"
) +
  xlab("Plant size") + ylab(NULL) +
  scale_x_continuous(breaks = c(-4, -2, 0, 2, 4)) +
  panel_margin+
    theme(
      axis.title.y = element_blank(),   
      axis.text.y  = element_blank(),   
      axis.ticks.y = element_blank(),   
      axis.line.y  = element_blank(),   
      panel.grid.major.y = element_blank(), 
      panel.grid.minor.y = element_blank())


## 图5(d)
fig5_d_panel <- plot_trait_year_raw(
  model_raw   = lme_full,
  data        = total_scaled_tda1,
  trait       = "Above_PC3",
  time_vals   = gt_vals,
  trait_range = trait_ranges$Above_PC3,
  title       = "(e)"
) +
  xlab("Leaf area") + ylab(NULL) +
  scale_x_continuous(breaks = c(-4, -2, 0, 2, 4)) +
  panel_margin+
    theme(
      axis.title.y = element_blank(),   
      axis.text.y  = element_blank(),   
      axis.ticks.y = element_blank(),   
      axis.line.y  = element_blank(),   
      panel.grid.major.y = element_blank(), 
      panel.grid.minor.y = element_blank())


## 图5(e)
fig5_e_panel <- plot_trait_year_raw(
  model_raw   = lme_full,
  data        = total_scaled_tda1,
  trait       = "Below_PC1",
  time_vals   = gt_vals,
  trait_range = trait_ranges$Below_PC1,
  title       = "(f)"
) +
  xlab("Root collaboration") +
  ylab(expression("Relative growth rate (RGR, log cm yr"^{-1}*")")) +
  scale_x_continuous(breaks = c(-4, -2, 0, 2, 4)) +
  panel_margin

## 图5(f)
fig5_f_panel <- plot_trait_year_raw(
  model_raw   = lme_full,
  data        = total_scaled_tda1,
  trait       = "Below_PC2",
  time_vals   = gt_vals,
  trait_range = trait_ranges$Below_PC2,
  title       = "(g)") +
  xlab("Root economics") + ylab(NULL) +
  scale_x_continuous(breaks = c(-4, -2, 0, 2, 4)) +
  panel_margin+
    theme(
      axis.title.y = element_blank(),   
      axis.text.y  = element_blank(),   
      axis.ticks.y = element_blank(),   
      axis.line.y  = element_blank(),   
      panel.grid.major.y = element_blank(), 
      panel.grid.minor.y = element_blank())

## ============================================================
## 3) 右侧下半部分：图6(c) 从原始函数开始重建
##    依赖对象：lme_full, total_scaled_tda1
## ============================================================

plot_3way_time2 <- function(model,
                            data,
                            x,
                            mod,
                            time = "growth_time",
                            time_vals = c(-1.7059989, -0.2379998, 1.2299992),
                            time_labels = c("1st year", "2nd year", "3rd year"),
                            mod_vals_num = c(-1, 0, 1),
                            mod_labels_num = c("Low", "Medium", "High"),
                            mod_vals_fac = NULL,
                            mod_labels_fac = NULL,
                            x_label = NULL,
                            y_label = "Predicted RGR",
                            title = NULL,
                            palette = c("#B7B2D0", "#E0A8A8", "#448870"),
                            ribbon_alpha = 0.18,
                            response_raw = "RGR",
                            show_points = TRUE,
                            point_alpha = 0.10,
                            point_size = 0.05,
                            points_by_mod = FALSE,
                            time_round_digits = 6) {

  stopifnot(x %in% names(data), mod %in% names(data), time %in% names(data))
  stopifnot(response_raw %in% names(data))

  if (is.null(x_label)) x_label <- x
  if (is.null(title)) title <- sprintf("(%s × %s) × %s", x, mod, time)

  mod_is_factor <- is.factor(data[[mod]]) || is.character(data[[mod]])

  if (!mod_is_factor) {
    term_mod <- sprintf("%s [%s]", mod, paste(mod_vals_num, collapse = ", "))
    group_levels <- as.character(mod_vals_num)
    group_labels <- mod_labels_num
    pal <- palette
  } else {
    if (is.null(mod_vals_fac)) mod_vals_fac <- levels(as.factor(data[[mod]]))
    if (is.null(mod_labels_fac)) mod_labels_fac <- mod_vals_fac

    term_mod <- sprintf("%s [%s]", mod, paste(mod_vals_fac, collapse = ", "))
    group_levels <- mod_vals_fac
    group_labels <- mod_labels_fac
    pal <- palette[seq_len(length(mod_vals_fac))]
  }

  term_time <- sprintf("%s [%s]", time, paste(time_vals, collapse = ", "))

  x_range <- as.numeric(quantile(data[[x]], probs = c(0.05, 0.95), na.rm = TRUE))

  pred <- ggpredict(model,
                    terms = c(x, term_mod, term_time),
                    type = "fixed")

  df <- as.data.frame(pred) %>%
    mutate(
      Mod  = factor(group, levels = group_levels, labels = group_labels),
      Year = factor(as.character(facet),
                    levels = as.character(time_vals),
                    labels = time_labels)
    )

  year_map <- tibble(
    gt = round(as.numeric(time_vals), time_round_digits),
    Year = factor(gt,
                  levels = round(as.numeric(time_vals), time_round_digits),
                  labels = time_labels)
  )

  data_pt <- data %>%
    mutate(gt = round(as.numeric(.data[[time]]), time_round_digits)) %>%
    left_join(year_map, by = "gt") %>%
    filter(!is.na(Year))

  p <- ggplot()

  if (show_points) {
    p <- p #+
      # geom_point(
      #   data = data_pt,
      #   aes(x = .data[[x]], y = .data[[response_raw]]),
      #   colour = "grey60",
      #   alpha  = point_alpha,
      #   size   = point_size
      # )
  }

  p <- p +
    geom_line(
      data = df,
      aes(x = x, y = predicted, colour = Mod),
      linewidth = 1.1
    ) +
    facet_wrap(~Year, nrow = 1) +
    scale_color_manual(values = pal) +
    scale_fill_manual(values = pal) +
    labs(x = x_label, y = y_label, colour = mod, fill = mod, title = title) +
    coord_cartesian(xlim = x_range) +
    theme_Publication() +
    theme(
      strip.background = element_rect(fill = "white", colour = NA),
      strip.text       = element_text(colour = "black"),
      legend.position  = "bottom"
    )

  p
}

fig6_c_panel <- plot_3way_time2(
  model = lme_full,
  data  = total_scaled_tda1,
  x = "Above_PC1",
  mod = "Below_PC2",
  mod_vals_num = c(-1.5, 0, 1.5),
  mod_labels_num = c("low", "Medium", "high plant size"),
  x_label = "Leaf economics (Above_PC1)",
  title = "(h) Leaf economics × Root economics"
) +
  ylim(c(0, 0.75)) +
  panel_margin

## ============================================================
## 4) patchwork 拼接
##    左列两行；右列三行（25%,25%,50%）
## ============================================================

left_col <- fig4_a_panel / fig4_b_panel +
  plot_layout(heights = c(1, 1))

design_right <- "
ABC
DEF
GGG
"

## 这里把第三排包起来，避免 p_tt1 这类分面图再次触发 wrap_dims
right_col <- wrap_plots(
  A = fig5_b_panel,
  B = fig5_c_panel,
  C = fig5_d_panel,
  D = fig5_e_panel,
  E = fig5_f_panel,
  F = plot_spacer(),
  G = wrap_elements(full = ggplotGrob(fig6_c_panel)),
  design = design_right
) +
  plot_layout(heights = c(1, 1, 2.3))   # 25%, 25%, 50%

fig_recombined_final <- left_col | right_col +
  plot_layout(widths = c(1, 1))

fig_recombined_final


out_dir <- "02_output_results_new/figs"
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)

ggsave(
  filename = file.path(out_dir, "new_figure5_combined.png"),
  plot     = fig_recombined_final,
  width    = 8,
  height   = 7,
  units    = "in",
  bg       = "white"
)

topptx(
 fig_recombined_final,
  filename = file.path(out_dir, "new_figure5_combined.pptx"),
  width = 8,
  height = 6
)
```


# New Figure 5
```{r}
library(patchwork)
library(ggplot2)

## ============================================================
## 1) 指定子图对象
## ============================================================

# 来自旧 Figure 4
new5_a_panel <- p_d   # 旧图4(d) -> 新图5(a)
new5_b_panel <- p_e   # 旧图4(e) -> 新图5(b)
new5_c_panel <- p_c   # 旧图4(c) -> 新图5(c)

# 来自旧 Figure 5
new5_d_panel <- p_light       # 旧图5(g)
new5_e_panel <- p_moist       # 旧图5(h)
new5_f_panel <- p_ap   # 旧图5(f)

# 来自旧 Figure 6
new5_g_panel <- p_bpc2_m      # 旧图6(d)
new5_h_panel <- p_bpc1_l      # 旧图6(e)

## ============================================================
## 2) 图例设置
##    建议只保留右侧最后需要的图例，其他先关掉
## ============================================================

new5_a_panel <- new5_a_panel + theme(legend.position = "none")
new5_b_panel <- new5_b_panel + theme(legend.position = "none")
new5_c_panel <- new5_c_panel + theme(legend.position = "none")

new5_d_panel <- new5_d_panel + theme(legend.position = "none")
new5_e_panel <- new5_e_panel + theme(legend.position = "none")
new5_f_panel <- new5_f_panel + theme(legend.position = "none")

new5_g_panel <- new5_g_panel + theme(legend.position = "none")
new5_h_panel <- new5_h_panel + theme(legend.position = "bottom")

## ============================================================
## 3) 左侧竖列：新图5(a)(b)(c)
## ============================================================

left_col <- new5_a_panel / new5_b_panel / new5_c_panel +
  plot_layout(heights = c(1, 1, 1))

## ============================================================
## 4) 右侧第一排：旧图5(g)(h)(f)
## ============================================================

right_top <- new5_d_panel | new5_e_panel | new5_f_panel

## ============================================================
## 5) 右侧第二排、第三排：旧图6(d)、旧图6(e)
##    这两个面板横跨整排
## ============================================================

right_col <- right_top / new5_g_panel / new5_h_panel +
  plot_layout(heights = c(1, 1.1, 1.1))

## ============================================================
## 6) 左右拼接
## ============================================================

fig5_new <- left_col | right_col +
  plot_layout(widths = c(1, 2.8))

fig5_new

library(ggplot2)
library(eoffice)

out_dir <- "02_output_results_new/figs"
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)

ggsave(
  filename = file.path(out_dir, "new_figure5_combined.png"),
  plot     = fig5_new,
  width    = 8,
  height   = 9,
  units    = "in",
  bg       = "white"
)

topptx(
  fig5_new,
  filename = file.path(out_dir, "new_figure5_combined.pptx"),
  width = 8,
  height = 9
)
```

```{r,eval=FALSE}

p_tt2 <- plot_3way_time2(
  model = lme_full,
  data  = total_scaled_tda1,
  x = "Above_PC3",
  mod = "Below_PC2",
  mod_vals_num = c(-2, 0, 2),
  mod_labels_num = c("low", "Medium", "high root econ"),
  x_label = "Leaf area (Above_PC3)",
  title = "(b) Leaf area × Root economics "
)+ylim(c(0,1.2))
p_tt2

p_apc2_light <- plot_3way_time2(
  model = lme_full,
  data  = total_scaled_tda1,
  x = "Above_PC2",
  mod = "Light",
  mod_vals_num = c(-1.5, 0, 1.5),
  mod_labels_num = c("Low", "Medium", "High light"),
  x_label = "Plant size (Above_PC2)",
  title = "(b) Plant size × Light"
)+ylim(c(0,1.2))
p_apc2_light

p_apc1_ap <- plot_3way_time2(
  model = lme_full,
  data  = total_scaled_tda1,
  x = "Above_PC1",
  mod = "AP",
  mod_vals_num = c(-1.5, 0, 1.5),
  mod_labels_num = c("Low", "Medium", "High light"),
  x_label = "Plant size (Above_PC2)",
  title = "(b) Plant size × Light")+ylim(c(0,1.2))
p_apc1_ap

p_compete <- plot_3way_time2(
  model = lme_full,
  data  = total_scaled_tda1,
  x = "Above_PC3",
  mod = "Compete",
  mod_vals_fac   = c("alone", "inter"),
  mod_labels_fac = c("Competition-free", "Competition"),
  x_label = "Leaf area (Above_PC3)",
  title = "(c) Leaf area × Competition",
  palette = c("#B7B2D0", "#e3a561", "#448870"),
)+ylim(c(0,1.1))
p_compete


```

### Figure 6 这部分基于部分生长模型的可视化结果
```{r}
library(lmerTest)
library(sjPlot)
library(ggeffects)
library(ggplot2)
library(patchwork)
library(dplyr)

col_levels <- c("#B7B2D0", "#e3a561", "#448870")

x_min_med_max <- function(x) {
  x <- x[is.finite(x)]
  c(min(x), median(x), max(x))
}

make_level_labels <- function(vals, prefix = c("Low", "Median", "High"), digits = 2) {
  sprintf("%s (%.2f)", prefix, round(as.numeric(vals), digits))
}

fit_lmm <- function(formula, dat) {
  lmer(
    formula,
    data = dat,
    REML = FALSE,
    control = lmerControl(optimizer = "bobyqa")
  )
}

pred_term <- function(var, vals) {
  sprintf("%s [%.6f, %.6f, %.6f]", var, vals[1], vals[2], vals[3])
}

plot_pred_lines <- function(pred_df, title, xlab, ylab, legend_title, legend_labels) {
  ggplot(pred_df, aes(x = x, y = predicted, color = group, fill = group)) +
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.18, color = NA) +
    geom_line(linewidth = 1.1) +
    scale_color_manual(values = col_levels, labels = legend_labels) +
    scale_fill_manual(values = col_levels, labels = legend_labels) +
    labs(title = title, x = xlab, y = ylab, color = legend_title, fill = legend_title) +
    theme_classic(base_size = 13) +
    theme(plot.title = element_text(face = "bold"))
}

combine_ab <- function(p_left, p_right, legend_pos = "bottom") {
  (p_left | p_right) +
    plot_layout(guides = "collect") &
    theme(legend.position = legend_pos)
}

years <- list("1st" = tda_1st,"2nd"=tda_2nd, "3rd" = tda_3rd)

# =========================
# 0) Model tables (optional)
# =========================
formula_moist <- RGR_scale ~ (Above_PC1 + Below_PC2) * Moisture + Above_PC1:Below_PC2 + (1|sp) + (1|Nenv)
mod_moist_1st <- fit_lmm(formula_moist, years[["1st"]])
mod_moist_2nd <- fit_lmm(formula_moist, years[["2nd"]])
mod_moist_3rd <- fit_lmm(formula_moist, years[["3rd"]])
tab_model(mod_moist_1st, mod_moist_3rd, title = "Moisture model")

# =========================
# 1) Above × Below (fix Moisture at mean)
# =========================
above_vals_shared <- x_min_med_max(c(years[["1st"]]$Above_PC1, years[["3rd"]]$Above_PC1))
lab_above <- make_level_labels(above_vals_shared)

pred_ab_below_1st <- ggpredict(
  mod_moist_1st,
  terms = c(
    "Below_PC2 [all]",
    pred_term("Above_PC1", above_vals_shared),
    "Moisture [mean]"
  )
)

pred_ab_below_2nd<- ggpredict(
  mod_moist_2nd,
  terms = c(
    "Below_PC2 [all]",
    pred_term("Above_PC1", above_vals_shared),
    "Moisture [mean]"
  )
)

p_int_above_below_1st <- plot_pred_lines(
  pred_ab_below_1st,
  title = "(a) 1st year",
  xlab = "Below_PC2",
  ylab = "Predicted RGR (scaled)",
  legend_title = "Above_PC1",
  legend_labels = lab_above
)+theme_Publication(base_size = 11)

p_int_above_below_2nd <- plot_pred_lines(
  pred_ab_below_2nd,
  title = "(b) 3rd year",
  xlab = "Below_PC2",
  ylab = "Predicted RGR (scaled)",
  legend_title = "Above_PC1",
  legend_labels = lab_above
)+theme_Publication(base_size = 11)

fig_int_above_below <- combine_ab(p_int_above_below_1st, p_int_above_below_2nd)
fig_int_above_below

# =========================
# 2) Above × Moisture (fix Below at median)
# =========================
formula_ap <- RGR_scale ~ (Above_PC1 + Below_PC2) * AP + Above_PC1:Below_PC2 + (1|sp) + (1|Nenv)
mod_ap_1st <- fit_lmm(formula_ap, years[["1st"]])
mod_ap_3rd <- fit_lmm(formula_ap, years[["3rd"]])
tab_model(mod_ap_1st, mod_ap_3rd, title = "AP model")

moist_vals_shared <- x_min_med_max(c(years[["1st"]]$Moisture, years[["3rd"]]$Moisture))
lab_moist <- make_level_labels(moist_vals_shared)

pred_ab_moist_1st <- ggpredict(
  mod_moist_1st,
  terms = c(
    "Above_PC1 [all]",
    pred_term("Moisture", moist_vals_shared),
    "Below_PC2 [median]"
  )
)

pred_ab_moist_3rd <- ggpredict(
  mod_moist_3rd,
  terms = c(
    "Above_PC1 [all]",
    pred_term("Moisture", moist_vals_shared),
    "Below_PC2 [median]"
  )
)

p_int_above_moist_1st <- plot_pred_lines(
  pred_ab_moist_1st,
  title = "(a) 1st year",
  xlab = "Above_PC1",
  ylab = "Predicted RGR (scaled)",
  legend_title = "Moisture",
  legend_labels = lab_moist
)+theme_Publication(base_size = 11)

p_int_above_moist_3rd <- plot_pred_lines(
  pred_ab_moist_3rd,
  title = "(b) 3rd year",
  xlab = "Above_PC1",
  ylab = "Predicted RGR (scaled)",
  legend_title = "Moisture",
  legend_labels = lab_moist
)+theme_Publication(base_size = 11)

fig_int_above_moist <- combine_ab(p_int_above_moist_1st, p_int_above_moist_3rd)
fig_int_above_moist 
# =========================
# 3) Below × Moisture (fix Above at median)
# =========================
pred_below_moist_1st <- ggpredict(
  mod_moist_1st,
  terms = c(
    "Below_PC2 [all]",
    pred_term("Moisture", moist_vals_shared),
    "Above_PC1 [median]"
  )
)

pred_below_moist_3rd <- ggpredict(
  mod_moist_3rd,
  terms = c(
    "Below_PC2 [all]",
    pred_term("Moisture", moist_vals_shared),
    "Above_PC1 [median]"
  )
)

p_int_below_moist_1st <- plot_pred_lines(
  pred_below_moist_1st,
  title = "(a) 1st year",
  xlab = "Below_PC2",
  ylab = "Predicted RGR (scaled)",
  legend_title = "Moisture",
  legend_labels = lab_moist)+theme_Publication(base_size = 11)

p_int_below_moist_3rd <- plot_pred_lines(
  pred_below_moist_3rd,
  title = "(b) 3rd year",
  xlab = "Below_PC2",
  ylab = "Predicted RGR (scaled)",
  legend_title = "Moisture",
  legend_labels = lab_moist)+theme_Publication(base_size = 11)

fig_int_below_moist <- combine_ab(p_int_below_moist_1st, p_int_below_moist_3rd)
fig_int_below_moist

# =========================
# 4) Above × AP (fix Below at median)
# =========================
ap_vals_shared <- x_min_med_max(c(years[["1st"]]$AP, years[["3rd"]]$AP))
lab_ap <- make_level_labels(ap_vals_shared)

pred_ab_ap_1st <- ggpredict(
  mod_ap_1st,
  terms = c(
    "Above_PC1 [all]",
    pred_term("AP", ap_vals_shared),
    "Below_PC2 [median]"
  )
)

pred_ab_ap_3rd <- ggpredict(
  mod_ap_3rd,
  terms = c(
    "Above_PC1 [all]",
    pred_term("AP", ap_vals_shared),
    "Below_PC2 [median]"
  )
)

p_int_above_ap_1st <- plot_pred_lines(
  pred_ab_ap_1st,
  title = "(a) 1st year",
  xlab = "Above_PC1",
  ylab = "Predicted RGR (scaled)",
  legend_title = "AP",
  legend_labels = lab_ap
)+theme_Publication(base_size = 11)

p_int_above_ap_3rd <- plot_pred_lines(
  pred_ab_ap_3rd,
  title = "(b) 3rd year",
  xlab = "Above_PC1",
  ylab = "Predicted RGR (scaled)",
  legend_title = "AP",
  legend_labels = lab_ap
)+theme_Publication(base_size = 11)

fig_int_above_ap <- combine_ab(p_int_above_ap_1st, p_int_above_ap_3rd)
fig_int_above_ap
# =========================
# 5) Below × AP (fix Above at median)
# =========================
pred_below_ap_1st <- ggpredict(
  mod_ap_1st,
  terms = c(
    "Below_PC2 [all]",
    pred_term("AP", ap_vals_shared),
    "Above_PC1 [median]"
  )
)

pred_below_ap_3rd <- ggpredict(
  mod_ap_3rd,
  terms = c(
    "Below_PC2 [all]",
    pred_term("AP", ap_vals_shared),
    "Above_PC1 [median]"
  )
)

p_int_below_ap_1st <- plot_pred_lines(
  pred_below_ap_1st,
  title = "(a) 1st year",
  xlab = "Below_PC2",
  ylab = "Predicted RGR (scaled)",
  legend_title = "AP",
  legend_labels = lab_ap
)+theme_Publication(base_size = 11)

p_int_below_ap_3rd <- plot_pred_lines(
  pred_below_ap_3rd,
  title = "(b) 3rd year",
  xlab = "Below_PC2",
  ylab = "Predicted RGR (scaled)",
  legend_title = "AP",
  legend_labels = lab_ap
)+theme_Publication(base_size = 11)

fig_int_below_ap <- combine_ab(p_int_below_ap_1st, p_int_below_ap_3rd)
fig_int_below_ap
# =========================
# Print all figures
# =========================


allfigs <- allfigs <- wrap_plots(
  p_int_above_below_1st + ggtitle("(a)")+theme(legend.position = "none"), 
  p_int_above_moist_1st + ggtitle("(b)")+theme(legend.position = "none"), 
  p_int_below_moist_1st + ggtitle("(c)")+theme(legend.position = "none"), 
  p_int_above_below_2nd + ggtitle("(d)"),
  p_int_above_moist_3rd + ggtitle("(e)"),
  p_int_below_moist_3rd + ggtitle("(f)"),
  ncol = 3
)
allfigs
ggsave(allfigs,file="02_output_results_new/figs/fig6_TE_TT.pdf",height = 6,width = 9)

topptx(allfigs,file="02_output_results_new/figs/fig6_TE_TT.pptx", width = 9, height = 6)


print(fig_int_above_below)   # Fig: Above_PC1 × Below_PC2 (Moisture fixed)
print(fig_int_above_moist)   # Fig: Above_PC1 × Moisture (Below fixed)
print(fig_int_below_moist)   # Fig: Below_PC2 × Moisture (Above fixed)
print(fig_int_above_ap)      # Fig: Above_PC1 × AP (Below fixed)
print(fig_int_below_ap)      # Fig: Below_PC2 × AP (Above fixed)


# Assumes you have already created:
# fig_int_above_below, fig_int_above_moist, fig_int_below_moist, fig_int_above_ap, fig_int_below_ap

library(patchwork)

all_interaction_figs <- (
  fig_int_above_below /
  fig_int_above_moist /
  fig_int_below_moist 
) +
  plot_layout(heights = rep(1, 3), guides = "collect") &
  theme(legend.position = "bottom")

print(all_interaction_figs)
```


### Figure SX. simple linear model between PC axis and RGR
```{r}
indtraits_tda_long <- indtraits_tda %>%
  pivot_longer(
    cols = all_of(trait.names),
    names_to = "trait",
    values_to = "values"
  ) %>%
  drop_na(values) %>%
  mutate(
    comp_year = factor(comp_year, levels = c("1st", "2nd", "3rd")),
    Nenv = factor(Nenv, levels = c("6", "5", "4", "9", "2", "3", "7", "8", "1")),
    trait = factor(trait, levels = trait.names)
  )

table(indtraits_tda$comp_year)

############################### Data Transformation for Visualization ###############################
# Reshape data to long format for PCA traits
pca_tda_long <- gather(pcatraits_tda[, c(
  "sp", "sp_abbr", "Nenv", "ring.num", "Light", "Moisture", "AP", "Compete", "growth_time", "final_height",
  "comp_year", "BM", "ABM", "RSR", "UBM", "RGR", "Above_PC1", "Above_PC2","Above_PC3", "Below_PC1", "Below_PC2")], trait, values, pca_traits)
pca_tda_long <- na.omit(pca_tda_long)

fig2c_pcatrait_com_sp_heter <- ggplot(pca_tda_long, aes(x = values, y = RGR)) +
  geom_point(aes(col = comp_year), alpha = 0.3, size = 0.1) +
  facet_grid(sp ~ trait, scales = "free") +
  geom_smooth(
    method = "lm",
    aes(group = comp_year, fill = comp_year),
    col = "black", alpha = 0.7, size = 0.15, se = TRUE
  ) +
  labs(
    x = "Standardized trait values",
    y = expression("Relative growth rate" ~ "(" * RGR * "," ~ log ~ cm ~ yr^{-1} * ")")
  ) +
  scale_color_manual(
    values = c("#B7B2D0", "#e3a561", "#448870"),
    labels = c("alone" = "Competition-free", "inter" = "Competition")
  ) +
  scale_fill_manual(
    values = c("#B7B2D0", "#e3a561", "#448870"),
    labels = c("alone" = "Competition-free", "inter" = "Competition")
  ) +
  scale_x_continuous(breaks = pretty(pca_tda_long$values, n = 5)) +
  ylim(c(0, 1.2)) +
  theme_Publication(base_size = 12) +scale_y_continuous(
  labels = label_number(accuracy = 0.1, trim = TRUE)) + 
  theme(
    legend.position = "bottom",
    legend.key.size = unit(0.4, "cm"),
    legend.text = element_text(size = 15),
    legend.title = element_blank(),
    axis.text.x = element_text(size = 10, colour = "black"),
    axis.text.y = element_text(size = 10, colour = "black")
  )

fig2c_pcatrait_com_sp_heter
ggsave(
  filename = "02_output_results_new/figs/figS10_pcatrait_com_sp_heter.pdf",
  plot = fig2c_pcatrait_com_sp_heter,
  width = 8,
  height = 9)
```

